---
title: "Figure_1_Plots"
author: "Jannik Ruwisch"
date: "2025-01-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      fig.width = 14,
                      fig.height = 7,
                      #dpi = 1200,
                      dev = "png",
                      cache = F,
                      fig.path= "C:/Users/Jannik/OneDrive/PNE_MHH/AG_Prasse_OneDrive/PPFE/PPFE_Nuc_Seq/Batch_integration/ppfe_epithelium/")
```

```{r}
library(SCP)
library(Seurat)
library(SeuratObject)
#library(renv)
#library(SCP)
library(dplyr)
library(tidyverse)
library(tidyseurat)
library(ggpubr)
```

```{r}
test <- readRDS("PPFE_cleaned_datasetes_ALL_RPCA_clean_integrated_22_12_2024.rds")
```


```{r}
# Color Palettes"
dis_pal <- palette_scp(x = test$disease.ident, palcolor = c("royalblue", "red3", "#ffd59b"))
cohort_pal_heat <- palette_scp(x = test$Cohort, palcolor = c("grey88", "grey58", "grey18"))
cohort_pal <-  c("royalblue", "red1", "darkred", "#ffd59b")
lin_pal <- palette_scp(x = test$Lineage,palcolor = c( "#cc79a7", "#009e73", "#5654e9", "#d55e00"  , "#e69f00"))
subject_pal <-  palette_scp(palette = "Set1", n = length(test$orig.ident %>% unique()), matched = F)
```

```{r}
folder <- choose.dir()
```


```{r}
# Start from here...
test <- test %>%
  mutate(Cohort = case_when(orig.ident_final == "PPFE040" ~ "Germany",
                            .default = Cohort))


test <- test %>%
  mutate(orig.ident = case_when(Project == "PPFE" ~ orig.ident,
                                  Project == "IPF" ~ substr(orig.ident, start = 1, stop = 10))
  )

#test$disCohort_FINAL %>% unique()

test$disCohort_FINAL <- paste0(test$Cohort, "_", test$disease.ident)
```



```{r}
# Adjust for patient idents based on encryptiont table 
test <- test%>%
  mutate(orig.ident_final = case_when(orig.ident == "CTRL_GLE822" ~ "CTR008",
                                      orig.ident == "CTRL_GLE891" ~ "CTR011",
                                      orig.ident == "CTRL_CTRL002" ~ "CTR002",
                                      orig.ident == "CTRL_GLR632" ~ "CTR072",
                                      orig.ident == "CTRL_GLR635" ~ "CTR073",
                                      orig.ident == "CTRL_CTRL007" ~ "CTR007",
                                      orig.ident == "CTRL_GLE639" ~ "CTR075",
                                      orig.ident == "CTRL_GLR609" ~ "CTR031",
                                      orig.ident == "CTRL_GLR725" ~ "CTR064",
                                      orig.ident == "CTRL_CTR048" ~ "CTR048",
                                      orig.ident == "CTRL_CTR068" ~ "CTR068",
                                      orig.ident == "CTRL_CTR051" ~ "CTR051",
                                      orig.ident == "CTRL_CTR071" ~ "CTR071",
                                      orig.ident == "CTRL_CTR058" ~ "CTR058",
                                      orig.ident == "CTRL_CTR069" ~ "CTR069",
                                      orig.ident == "CTRL_CTR070" ~ "CTR070",
                                      orig.ident == "PPFE_PPFE015a" ~ "PPFE015",
                                      orig.ident == "PPFE_PPFE015b" ~ "PPFE015",
                                      orig.ident == "PPFE_PPFE005" ~ "PPFE002",
                                      orig.ident == "PPFE_PPFE42" ~ "PPFE006",
                                      
                                      orig.ident == "PPFE_PPFE001" ~ "PPFE001",
                                      orig.ident == "PPFE_PPFE002" ~ "PPFE002",
                                      orig.ident == "PPFE_PPFE003" ~ "PPFE003",
                                      orig.ident == "PPFE_PPFE004" ~ "PPFE004",
                                      orig.ident == "PPFE_PPFE005" ~ "PPFE005",
                                      orig.ident == "PPFE_PPFE006" ~ "PPFE006",
                                      orig.ident == "PPFE_PPFE007" ~ "PPFE007",
                                      orig.ident == "PPFE_PPFE008" ~ "PPFE008",
                                      orig.ident == "PPFE_PPFE009" ~ "PPFE009",
                                      orig.ident == "PPFE_PPFE010" ~ "PPFE010",
                                      orig.ident == "PPFE_PPFE011" ~ "PPFE011",
                                      orig.ident == "PPFE_PPFE012" ~ "PPFE012",
                                      orig.ident == "PPFE_PPFE013" ~ "PPFE013",
                                      orig.ident == "PPFE_PPFE014" ~ "PPFE014",
                                      orig.ident == "PPFE_PPFE016" ~ "PPFE016",
                                      orig.ident == "PPFE_PPFE017" ~ "PPFE017",
                                      orig.ident == "PPFE_PPFE018" ~ "PPFE018",
                                      orig.ident == "PPFE_PPFE019" ~ "PPFE019",
                                      orig.ident == "PPFE_PPFE020" ~ "PPFE020",
                                      orig.ident == "PPFE_PPFE21" ~ "PPFE021",
                                      orig.ident == "PPFE_PPFE22" ~ "PPFE022",
                                      orig.ident == "PPFE_PPFE23" ~ "PPFE023",
                                      orig.ident == "PPFE_PPFE24" ~ "PPFE024",
                                      orig.ident == "PPFE_PPFE25" ~ "PPFE025",
                                      orig.ident == "PPFE_PPFE26" ~ "PPFE026",
                                      orig.ident == "PPFE_PPFE27" ~ "PPFE027",
                                      orig.ident == "PPFE_PPFE28" ~ "PPFE028",
                                      orig.ident == "PPFE_PPFE29" ~ "PPFE029",
                                      orig.ident == "PPFE_PPFE30" ~ "PPFE030",
                                      orig.ident == "PPFE_PPFE31" ~ "PPFE031",
                                      orig.ident == "PPFE_PPFE32" ~ "PPFE032",
                                      orig.ident == "PPFE_PPFE33" ~ "PPFE033",
                                      orig.ident == "PPFE_PPFE34" ~ "PPFE034",
                                      orig.ident == "PPFE_PPFE35" ~ "PPFE035",
                                      orig.ident == "PPFE_PPFE36" ~ "PPFE036",
                                      orig.ident == "PPFE_PPFE37" ~ "PPFE037",
                                      orig.ident == "PPFE_PPFE38" ~ "PPFE038",
                                      orig.ident == "PPFE_PPFE39" ~ "PPFE039",
                                      orig.ident == "PPFE_PPFE40" ~ "PPFE040",
                                      orig.ident == "PPFE_PPFE41" ~ "PPFE041",
                                      orig.ident == "PPFE_PPFE43" ~ "PPFE043",
                                      orig.ident == "PPFE_PPFE44" ~ "PPFE044",
                                      .default = orig.ident)
         )

test$orig.ident_final %>% unique()

test$orig.ident_old <- test$orig.ident
test$orig.ident <- test$orig.ident_final

```

```{r}
# Rename the final metadata labels
test$Celltype <- test$Celltype_ID_final


test <- test %>%
  mutate(Cohort = case_when(Project == "PPFE" ~ Cohort,
                                  Project == "IPF" ~ "Belgium")
       )


test <- test %>%
  mutate(disease_cohort = paste(disease.ident, Cohort, sep = "_"))
```

## save correct datasets
```{r}
test@meta.data %>% colnames()
metadata_to_keep <- c("orig.ident_final", "Celltype", "disease.ident", "Cohort", "disease_cohort", "Lineage")

CellDimPlot(test, group.by = metadata_to_keep, combine = F, raster = F, reduction = "ReUMAP_50")

# Remove metadata
test@meta.data <- test@meta.data[, metadata_to_keep]


# Remove embeddings besides pca and ReUMAP  
# Keep only the specified reductions
test@reductions <- test@reductions[c("pca", "ReUMAP_50")]

# Rename 'ReUMAP_18' to 'umap'
names(test@reductions)[names(test@reductions) == "ReUMAP_50"] <- "umap"

# Update the key for the new 'umap' reduction
test@reductions$umap@key <- "UMAP_"

# Check result
CellDimPlot(test, 
            group.by = "Celltype",
            combine = F,
            raster = F)
```
```{r}
# save the final seurat objects
folder = choose.dir()
saveRDS(test, file = file.path(folder, "PPFE_Merged_Lineage.rds"))


```



## Color Palette Settings
```{r}
pal_cluster_scp_pal <- c(
  # Epithelial Cells
  "#7F3C8D",  # AEC2
  "#199784",  # AEC1
  "#715588",  # Club
  "#8B8B60",  # AEC_intermediate
  "#A5AA99",  # Basal
  "#CC831E",  # Secretory
  "#ff084a",  # Aberrant_Basaloid | corrected
  "#446293",  # Goblet
  "#8BAC5C",  # Ciliated
  "#FDBF6F",  # PNEC
  "#FDBF6F",  # Mesothelium
  "#000000",  # Schwann_Neu (Not specified, placeholder)

  # Fibroblasts & Stromal Cells
  "#33A02C",  # Pericytes
  "#A6CEE3",  # Alveolar_Fb
  "#6A3D9A",  # Adventitial_Fb
  "#FB9A99",  # Subpleural_Fb
  "#CAB2D6",  # Adventitial_like_Fb
  "#B2DF8A",  # Smooth_Muscle
  "#E31A1C",  # CTHRC1_MyoFb
  "#1F78B4",  # Airway_Fb

  # Endothelial Cells
  "#4DBBD5",  # EC_pulmonary_venous
  "#F39B7F",  # EC_arterial
  "#3C5488",  # EC_general_cap
  "#E64B35",  # EC_systemic_venous
  "#8491B4",  # EC_aerocyte
  "#00A087",  # EC_lymphatic

  # Myeloid Cells
      "#BA4769", # Fibrotic_Macrophage"#C19D94", | Updated !
  "#BE777C",  # Monocyte_derived_Macrophage
  "#FF6F00",  # Alveolar_Macrophage
  "#BA7769",  # Monocyte_nonClassical
  "#6A7998",  # Monocyte_Classical
  "#1D5152",  # DC2
  "#94CABB",  # DC1
  "#CB1700",  # Mast
  "#B7BBC9",  # IFNresp_Monocyte_derived_Macrophage
  "#58644F",  # pDC
  "#3F4041",  # Langerhans-Cells
  "#67549A",  # Macrophage_Cycling
  "#217885",  # Neutro

  # Lymphoid Cells
  "#E6F5B7",  # CD4+_TC_naive
  "#D9D7C9",  # CD4+_TC_effector_memory
  "#D9B382",  # CD4+_TC_TH1 (Not explicitly named, assumed TH1 effector memory)
  "#8DD3C7",  # Treg
  "#DED7DA",  # NK_Cells_CD16+
  "#C7D98C",  # CD8+_TC_resident_memory
  "#FFED6F",  # B_Cells_memory
  "#D8C965",  # CD8+_TC_effector_memory
  "#88ADCC",  # ILC3
  "#F9CDE4",  # gamma_delta_TC
  "#D9B382",  # CD4+_TC_TH17 (Assumed from similar CD4 TC subsets)
  "#D9B382",  # CD4+_TC_Cytotoxic (Not explicitly defined, assumed)
  "#D6EBB2",  # B_Cells_naive
  "#C2ADC0",  # Plasma_Cells
  "#C69FC7" )  # NK_Cells_CD16-
  

  lin_pal <- palette_scp(x = test$Lineage,palcolor = c( "#cc79a7", "#009e73", "#5654e9", "#d55e00"  , "#e69f00"))
pal_cluster_scp_pal

pal_dis_pal <- palette_scp(x = test$disease.ident%>% unique(), palcolor = c("royalblue", "red3", "#ffd59b"), matched = F)
pal_dis_pal

pal_cohort_pal_heat <- palette_scp(x = c("Germany", "France",  "Belgium"), palcolor = c("grey88", "grey58", "grey18"), matched = F)
pal_cohort_pal_heat

pal_cohort_pal <-  c("royalblue", "red1", "darkred", "#ffd59b")
#lin_pal <- palette_scp(x = PPFE_endo$Lineage,palcolor = c( "#cc79a7", "#009e73", "#5654e9", "#d55e00"  , "#e69f00"))
pal_subject_pal <-  palette_scp(palette = "igv", x = test$orig.ident_final %>% unique(), matched = F) 
pal_subject_pal

```


# 1. Plot the Cellfrequencies
```{r}
Batch2_Fren_IDs
```

### Overall Dataset
```{r, fig.width=14, fig.height=7}
# Proportion / cell number composition per cluster
ggData = data.frame(prop.table(table(test$Lineage, test$orig.ident_final), margin = 2))
colnames(ggData) = c("Celltype_ID", "Orig.Ident",  "value")

ggData <- ggData %>%
 mutate(Disease = substr(Orig.Ident, 1,4))

subset(test, disCohort_FINAL == "France_PPFE")$orig.ident_final %>% unique()-> Batch2_Fren_IDs 
subset(test, disCohort_FINAL == "Germany_PPFE")$orig.ident_final %>% unique()-> BAtch1_Han_IDs 

BAtch1_Han_IDs <- c(BAtch1_Han_IDs, "PPFE040") # correction !! 
Batch2_Fren_IDs <- c("PPFE021", "PPFE022", "PPFE025", "PPFE026", "PPFE028", "PPFE029", 
                 "PPFE031", "PPFE032", "PPFE033", "PPFE034", "PPFE035", "PPFE037", 
                 "PPFE038",  "PPFE023", "PPFE024", "PPFE036", "PPFE039")#"PPFE040",) correction ! is German !!
subset(test, disCohort_FINAL == "Germany_CTRL")$orig.ident_final %>% unique()-> CTRL_ids
```


```{r, fig.width=14, fig.height=7}
ggData <- ggData %>%
 mutate(Disease = case_when(Orig.Ident %in% Batch2_Fren_IDs ~ "France_PPFE",
                            Orig.Ident %in% BAtch1_Han_IDs ~ "Germany_PPFE",
                            Orig.Ident %in% CTRL_ids ~ "Germany_CTRL",
                            .default = "Belgium_IPF")
        )

ggData$Disease <- factor(ggData$Disease, levels = c("Germany_CTRL", "France_PPFE", "Germany_PPFE", "Belgium_IPF"))



test$Celltype_ID_final_coarse %>% unique() -> Celltypes
levels(Celltypes)

#cluster_scp_pal <- SCP::palette_scp(x = c( "Epithelium","Mesenchyme","Endothelium", "Lymphoid","Myeloid"), palcolor = lin_pal)

patient_order <- ggData %>%
  dplyr::filter(Celltype_ID == 'Lymphoid') %>% 
  arrange(Disease, desc(value)) %>% # Zuerst nach Disease, dann nach value absteigend
  pull(Orig.Ident)

# Convert Patient_ID to a factor with the specific order
ggData$Orig.Ident <- factor(ggData$Orig.Ident, levels = patient_order)


# Calculate median values for each Celltype_ID within each Disease group
medians <- ggData %>%
  dplyr::filter(Disease =="Germany_PPFE") %>%
  group_by(Disease, Celltype_ID) %>%
  summarize(median_value = median(value)) %>%
  ungroup()

# Order Celltype_ID based on median values
ordered_celltypes <- medians %>%
  group_by(Celltype_ID) %>%
  summarize(overall_median = median(median_value)) %>%
  arrange(overall_median) %>%
  pull(Celltype_ID)
ordered_celltypes

# Update the levels of Celltype_ID in ggData
ggData$Celltype_ID <- factor(ggData$Celltype_ID, levels = c( "Epithelium","Mesenchyme","Endothelium", "Myeloid","Lymphoid")
                             )


# Plot with normalized bars
p <- ggplot(ggData, aes(x = Orig.Ident, y = value, fill = Celltype_ID)) +
  geom_bar(stat = "identity", position = "fill", size = 0.5, color = "black") +
  labs(title = "Stacked Bar Plot of Cell Type Frequencies by Patient (Ordered by Disease Cohort)",
       x = "Patient ID",
       y = "Proportion (%)") + 
  scale_fill_manual(values = lin_pal) +
  scale_color_manual(values = lin_pal) +
  theme_scp() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p
```

```{r, fig.width=30}
p <- ggplot(ggData, aes(x = Orig.Ident, y = value, fill = Celltype_ID)) +
  geom_bar(stat = "identity", position = "stack", size = 0.5, color = "black") +
  labs(#title = "Stacked Bar Plot of Cell Type Frequencies by Patient",
       x = NULL,
       y = "Frequency (%)") + 
  scale_fill_discrete(lin_pal) +
  scale_color_discrete(lin_pal) +
  scale_fill_manual(values = lin_pal) +
  scale_color_manual(values = lin_pal) +
  theme_scp() +
  theme(element_text(angle = 45, hjust = 1),
        axis.text.x.bottom = element_blank(), 
        axis.ticks.length.x = unit(0, "cm"))+
  facet_wrap(~Disease, scales = "free_x", ncol = 4) + NoLegend()+RotatedAxis()

p
```

```{r}
choose.dir() ->folder
```
```{r}
png((filename = file.path(folder,"Figure_1Stackedbarplots.png")), units = "in", width = 16, height = 6.5, res = 1200)
p
dev.off()

pdf((file = file.path(folder,"Figure_1Stackedbarplots.pdf")),width = 16, height = 6.5)
p
dev.off()
```

```{r, fig.width=24}
p <- ggplot(ggData, aes(x = Orig.Ident, y = value, fill = Celltype_ID)) +
  geom_bar(stat = "identity", position = "stack", size = 0.5, color = "black") +
  labs(title = "Stacked Bar Plot of Cell Type Frequencies by Patient",
       x = "Patient ID",
       y = "Frequency (%)") + 
  scale_fill_discrete(lin_pal) +
  scale_color_discrete(lin_pal) +
  scale_fill_manual(values = lin_pal) +
  scale_color_manual(values = lin_pal) +
  theme_scp() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))#+ facet_wrap(~Disease)
p
```

```{r}
png((filename = file.path(folder,"Figure_1Stackedbarplots_long.png")), units = "in", width = 18, height = 12, res = 600)
p
dev.off()
```

```{r}
# Convert disease.ident and Celltype to factors
ggData$Disease <- as.factor(ggData$Disease)
ggData$Celltype_ID <- as.factor(ggData$Celltype_ID)

# Calculate the median frequency for each cell type
median_frequencies <- ggData %>%
  group_by(Celltype_ID) %>%
  summarize(median_freq = median(value)) %>%
  arrange(-median_freq)

# Reorder the Celltype factor levels based on median frequency
ggData$Celltype_ID <- factor(ggData$Celltype_ID, levels = median_frequencies$Celltype_ID)

# Plot
p <- ggplot(ggData, aes(x = Celltype_ID, y = value, fill = Disease)) +
  geom_boxplot() +
  labs(title = "Comparison of Cell Type Frequencies Between Disease Groups",
       x = "Cell Type",
       y = "Frequency (%)") +
  scale_fill_manual(values = pal_cohort_pal) +
  scale_color_manual(values = pal_cohort_pal) +
  theme_scp() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Add statistical comparisons
p <- p + stat_compare_means(aes(group = Disease), method = "t.test", label = "p.signif")

#==============================================

# Plot
p <- ggplot(ggData, aes(x = Celltype_ID, y = value, fill = Disease)) +
  geom_boxplot(outlier.shape = NA) +
  labs(title = "Relative Endothelial Celltype Frequencies",
       x = "Cell Type",
       y = "Frequency (%)") +
  scale_fill_manual(values = pal_cohort_pal) +
  scale_color_manual(values = pal_cohort_pal) +
  theme_scp() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Add statistical comparisons
p + stat_compare_means(aes(group = Disease),# method = "t.test", 
                       label = "p.signif")


png(filename = file.path(folder,"Figure_1_Boxplots.png"), height = 7, width = 9, units = "in", res = 1200)
#pdf("PPFE_MES_UMAP.pdf", height = 7, width = 7)
p+ stat_compare_means(aes(group = Disease),# method = "t.test", 
                       label = "p.signif")
dev.off()

pdf(file.path(folder,"Figure_1_Boxplots.pdf"),  height = 7, width = 9)
p
dev.off()

#======================================
ggData_1 <- subset(ggData, Celltype_ID != "Alveolar_Fb" & 
                     Celltype_ID != "Smooth_Muscle" & 
                     Celltype_ID != "Adventitial_Fb" &
                     Celltype_ID != "CTHRC1_MyoFb" &
                     Celltype_ID != "Pericytes") 

# Calculate the median frequency for each cell type
median_frequencies <- ggData_1 %>%
  group_by(Celltype_ID) %>%
  summarize(median_freq = median(value)) %>%
  arrange(-median_freq)

# Reorder the Celltype factor levels based on median frequency
ggData_1$Celltype_ID <- factor(ggData_1$Celltype_ID, levels = median_frequencies$Celltype_ID)

# Plot
p <- ggplot(ggData_1, aes(x = Celltype_ID, y = value, fill = Disease)) +
  geom_boxplot(outlier.shape = NA
    ) +
  labs(title = NULL,
       x = "Cell Type",
       y = "Frequency (%)") +
  scale_fill_manual(values = pal_cohort_pal) +
  scale_color_manual(values =pal_cohort_pal) +
  theme_scp(aspect.ratio = 1) + 
  theme(axis.text.x = element_text(angle = 70, hjust = 1))#+ylim(0, 0.1)

# Add statistical comparisons
p1 <- p + stat_compare_means(aes(group = Disease),# method = "t.test", 
                            label = "p.signif")

# Display the plot
print(p1)

ggData_2 <- subset(ggData, Celltype_ID == "Alveolar_Fb" | 
                     Celltype_ID == "Smooth_Muscle" | 
                     Celltype_ID == "Adventitial_Fb" |
                     Celltype_ID == "CTHRC1_MyoFb" |
                     Celltype_ID == "Pericytes")

# Calculate the median frequency for each cell type
median_frequencies <- ggData_2 %>%
  group_by(Celltype_ID) %>%
  summarize(median_freq = median(value)) %>%
  arrange(-median_freq)

# Reorder the Celltype factor levels based on median frequency
ggData_2$Celltype_ID <- factor(ggData_2$Celltype_ID, levels = median_frequencies$Celltype_ID)

# Plot
p <- ggplot(ggData_2, aes(x = Celltype_ID, y = value, fill = Disease)) +
  geom_boxplot(outlier.shape = NA) +
  labs(title =NULL,
       x = "Cell Type",
       y = "Frequency (%)") +
 scale_fill_manual(values = pal_cohort_pal) +
  scale_color_manual(values = pal_cohort_pal) +
  theme_scp(aspect.ratio = 1) +
  theme(axis.text.x = element_text(angle = 70, hjust = 1))

# Add statistical comparisons
p2 <- p + stat_compare_means(aes(group = Disease), #method = "t.test", 
                            label = "p.signif")

# Display the plot
print(p2)

# Print together
p <- ggarrange(plotlist = list(p2, p1), ncol = 2, common.legend = T, align = "h")

png(filename = file.path(folder,"Figure_6_Boxplots_split.png"), height = 7, width = 9, units = "in", res = 1200)
#pdf("PPFE_MES_UMAP.pdf", height = 7, width = 7)
p
dev.off()

pdf(file.path(folder,"Figure_6_Boxplots_split.pdf"),  height = 7, width = 9)
p
dev.off()
```

## 3A) UMAP - Celltype_ID_final
```{r}
p <- CellDimPlot(test, group.by = "Celltype_ID_final", raster = F, pt.size = 0.15, reduction = "ReUMAP_50", palette = "Paired", show_stat = F, xlab = "UMAP_1", ylab = "UAMP_2", legend.position = "none")+
  theme(axis.text = element_blank(),   # Remove axis text
        axis.ticks = element_blank())  # Remove axis ticks

pdf("PPFE_ALL_UMAP.pdf", height = 14, width = 14)
p
dev.off()
p
```
## 3B) UMAP - Lineage
```{r}
p <-DimPlot(test, group.by = c("Lineage"),
          shuffle = T, raster = F, label = F, combine = T, pt.size = 0.15,
          reduction = "ReUMAP_50", 
        cols =lin_pal)+ 
  theme_scp()+
  labs(x = "UMAP_1", y = "UMAP_2")+
  #NoLegend()+
  ggtitle("Subject")+
  scale_color_manual(values = as.vector(lin_pal))+
  theme(axis.text = element_blank(), legend.direction = "horizontal",   # Remove axis text
        axis.ticks = element_blank(),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 25, face = "bold"),
        legend.key.size = unit(2, "cm"),
        legend.position = "bottom")+
  guides(color = guide_legend(override.aes = list(size = 16)))  
        
         # Remove axis ticks

pdf(file.path(folder, "PPFE_ALL_MERGED_UMAP_Lineage_legend.pdf"), height = 14, width = 14)
p
dev.off()
#p

```

## 3C) UMAP - Subject
```{r}
#SCP::palette_scp(palette = "Set1", n = length(test$orig.ident %>% unique()), matched = F) -> palette

p <- DimPlot(test, group.by = c("orig.ident"), 
             shuffle = T, 
             raster = F, 
             label = F, 
             combine = T, 
             pt.size = 0.15,
             reduction = "ReUMAP_50")+ 
  theme_scp()+
  labs(x = "UMAP_1", y = "UMAP_2")+
  #NoLegend()+
  ggtitle("Subject")+
  scale_color_manual(values = as.vector(pal_subject_pal))+
  theme(axis.text = element_blank(), legend.direction = "vertical",   # Remove axis text
        axis.ticks = element_blank(),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 25, face = "bold"),
        legend.key.size = unit(2, "cm"),
        legend.position = "bottom")+
  NoLegend()+
  guides(color = guide_legend(override.aes = list(size = 16)))  

pdf(file.path(folder, "PPFE_IPF_EPI_subject_UMAP_RPCA1.pdf"), height = 14, width = 14)
p
dev.off()
```
## 3d) UMAP - Disease

```{r}
p <- DimPlot(test, group.by = c("disease.ident"),
          shuffle = T, raster = F, label = F, combine = T, pt.size = 0.15,
          reduction = "ReUMAP_50", 
        cols =dis_pal)+
  theme_scp()+ggtitle("Disease")+#NoLegend()+
  labs(x = "UMAP_1", y = "UMAP_2")+
  theme(axis.text = element_blank(), 
        legend.direction = "horizontal",# Remove axis text
        axis.ticks = element_blank(),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 25, face = "bold"),
        #legend.key.size = unit(6, "cm"),
        legend.position = "bottom")+
  guides(color = guide_legend(override.aes = list(size = 16)))  


pdf(file.path(folder, "PPFE_IPF_Lineage_Disease_UMAP_RPCA1_legend.pdf"), height = 14, width = 14)
p
dev.off()
```


```{r}
png(filename = "PPFE_IPF_EPI_subject_Disease_UMAP_RPCA1.png", height = 14, width = 28, units = "in", res = 600)
#pdf(file = "PPFE_Mesenchyme_subject_Disease_UMAP.pdf", width = 14, height = 7)
p
dev.off()



```

## UMAP_Disease_Subject
```{r}
p <- DimPlot(test, group.by = c("orig.ident_final"),shuffle = T, raster = F, label = F, combine = T, reduction = "ReUMAPone", pt.size = 0.01, 
             cols = pal_subject_pal)+
  theme_scp()+NoLegend()+ggtitle("Subject")+
  theme(axis.text = element_blank(),   # Remove axis text
        axis.ticks = element_blank(), 
        legend.title =element_text(face = "bold")) +
          labs(x = "UMAP_1",  # New label for x-axis
               y = "UMAP_2"   # New label for y-axis
  )|
  DimPlot(test, group.by = c("disease.ident"),shuffle = T, raster = F, label = F, combine = T, reduction = "ReUMAPone", pt.size = 0.01, cols = pal_dis_pal)+
    theme_scp()+
  theme(axis.text = element_blank(),   # Remove axis text
        axis.ticks = element_blank(), 
        legend.title =element_text(face = "bold"))+
          labs(x = "UMAP_1",  # New label for x-axis
               y = "UMAP_2"   # New label for y-axis
  )+
  NoLegend()+ggtitle("Disease")
p
png(filename = file.path(folder, "Figure4_Disease_Cohort_UMAP.png"), height = 7, width = 14, units = "in", res = 1200)
#pdf(file = "PPFE_Mesenchyme_subject_Disease_UMAP.pdf", width = 14, height = 7)
p
dev.off()

pdf(file.path(folder, "Figure4_Disease_Cohort_UMAP.pdf"), height = 7, width = 14)
#pdf(file = "PPFE_Mesenchyme_subject_Disease_UMAP.pdf", width = 14, height = 7)
p
dev.off()


p_leg <- DimPlot(PPFE_mes, group.by = c("orig.ident_final"),shuffle = T, raster = F, label = F, combine = T, reduction = "ReUMAPone", pt.size = 0.01, 
             cols = pal_subject_pal)+
  theme_scp()+ggtitle("Subject")+
  theme(axis.text = element_blank(),   # Remove axis text
        axis.ticks = element_blank(), 
        legend.title =element_text(face = "bold")) +
          labs(x = "UMAP_1",  # New label for x-axis
               y = "UMAP_2"   # New label for y-axis
  )|
  DimPlot(PPFE_mes, group.by = c("disease.ident"),shuffle = T, raster = F, label = F, combine = T, reduction = "ReUMAPone", pt.size = 0.01, cols = pal_dis_pal)+
    theme_scp()+
  theme(axis.text = element_blank(),   # Remove axis text
        axis.ticks = element_blank(), 
        legend.title =element_text(face = "bold"))+
          labs(x = "UMAP_1",  # New label for x-axis
               y = "UMAP_2"   # New label for y-axis
  )+ggtitle("Disease")
p_leg

png(filename = file.path(folder, "Figure4_Disease_Cohort_UMAP_leg.png"), height = 7, width = 14, units = "in", res = 1200)
#pdf(file = "PPFE_Mesenchyme_subject_Disease_UMAP.pdf", width = 14, height = 7)
p_leg
dev.off()

pdf(file.path(folder, "Figure4_Disease_Cohort_UMAP_leg.pdf"), height = 7, width = 14)
#pdf(file = "PPFE_Mesenchyme_subject_Disease_UMAP.pdf", width = 14, height = 7)
p_leg
dev.off()
```


```{r, fig.width=28, fig.height=14}
png(filename = file.path(folder, "Figure6_Disease_Cohort_UMAP.png"), height = 7, width = 14, units = "in", res = 1200)
#pdf(file = "PPFE_Mesenchyme_subject_Disease_UMAP.pdf", width = 14, height = 7)
p
dev.off()

pdf(file.path(folder, "Figure6_Disease_Cohort_UMAP.pdf"), height = 7, width = 14)
#pdf(file = "PPFE_Mesenchyme_subject_Disease_UMAP.pdf", width = 14, height = 7)
p
dev.off()
p
```

## Statistics

#### Wilcox Rank Sum Test & Tukey-Test posthoc
```{r}
library(rstatix)
library(writexl)
summary_df <- ggData %>%
  group_by(Celltype_ID, Disease) %>%
  summarise(
    mean_value = mean(value, na.rm = TRUE),
    sd_value = sd(value, na.rm = TRUE)
  ) %>%
  ungroup()
write_xlsx(summary_df, file.path(folder, "PPFE_IPF_all_by_lineage_freq.xlsx"))

wilcox_results <- ggData %>%
  group_by(Celltype_ID) %>%
  wilcox_test(value ~ Disease) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj") %>%
  ungroup()

wilcox_results
write_xlsx(wilcox_results, file.path(folder, "PPFE_IPF_Mes_by_lineage_freq_wilcox.xlsx"))


final_df <- summary_df %>%
  left_join(wilcox_results, by = "Celltype_ID")# %>%
  #left_join(tukey_results, by = c("Celltype_ID", "group1", "group2"))
final_df

library(writexl)
write_xlsx(final_df, file.path(folder, "PPFE_IPF_allmerged_by_lineaeg_freq_Wilcox_rank_sum_test_merged.xlsx"))
```

```{r}
library(dplyr)
library(openxlsx)

# Initialize a workbook
wb <- createWorkbook()

# Germany_PPFE
median_frequencies_Germany_PPFE <- ggData %>%
  filter(Disease == "Germany_PPFE") %>%
  group_by(Celltype_ID) %>%
  summarize(
    median_freq = mean(value),
    sd_freq = sd(value)
  ) %>%
  arrange(-median_freq)

addWorksheet(wb, "Germany_PPFE")
writeData(wb, "Germany_PPFE", median_frequencies_Germany_PPFE)

# France_PPFE
median_frequencies_France_PPFE <- ggData %>%
  filter(Disease == "France_PPFE") %>%
  group_by(Celltype_ID) %>%
  summarize(
    median_freq = mean(value),
    sd_freq = sd(value)
  ) %>%
  arrange(-median_freq)

addWorksheet(wb, "France_PPFE")
writeData(wb, "France_PPFE", median_frequencies_France_PPFE)

# Germany_CTRL
median_frequencies_Germany_CTRL <- ggData %>%
  filter(Disease == "Germany_CTRL") %>%
  group_by(Celltype_ID) %>%
  summarize(
    median_freq = mean(value),
    sd_freq = sd(value)
  ) %>%
  arrange(-median_freq)

addWorksheet(wb, "Germany_CTRL")
writeData(wb, "Germany_CTRL", median_frequencies_Germany_CTRL)

# Dutch_IPF
median_frequencies_Dutch_IPF <- ggData %>%
  filter(Disease == "Durch_IPF") %>%
  group_by(Celltype_ID) %>%
  summarize(
    median_freq = mean(value),
    sd_freq = sd(value)
  ) %>%
  arrange(-median_freq)

addWorksheet(wb, "Durch_IPF")
writeData(wb, "Durch_IPF", median_frequencies_Dutch_IPF)


# Save the workbook
saveWorkbook(wb, "PPFE_IPF_Celltype_Frequencies_ALL_b1b2_RPCA.xlsx", overwrite = TRUE)


```




### Marker Gene Selection
```{r}
Lineage_genes <- c( "FXYD3", "EPCAM", "ELF3",#"EPCAM", "SLC34A2", "KRT7", "NKX2-1",
           "COL1A2", "DCN", "MFAP4", #"COL1A1", "COL3A1", "ELN", "COL6A2",
           "CLDN5", "CDH5", "CLEC14A", # "CDH5",  "PECAM1", "VWF", "BMPR2",
           "FCER1G", "C1orf162", "CLEC7A", #"CD68", "LYZ", "TLR2", "CSF1R",
           "CORO1A", "ISG20",  "PTPRC" #"CD2", "IL7R", "IL2RG", "CCL5"
           )
```

### ReOrder the Celltypes
```{r}
test$Celltype_ID_final_heat %>% unique()

test$Celltype_ID_final_heat <- factor(test$Celltype_ID_final, levels = c(
    "AEC2", "AEC1", "Club", "AEC_intermediate", "Basal", "Secretory", 
    "Aberrant_Basaloid", "Goblet", "Ciliated", "PNEC", "Mesothelium","Schwann_Neu",
   "Pericytes", "Alveolar_Fb", "Adventitial_Fb", "Subpleural_Fb", 
    "Adventitial_like_Fb", "Smooth_Muscle", "CTHRC1_MyoFb", "Airway_Fb",
   "EC_pulmonary_venous", "EC_arterial", "EC_general_cap", "EC_systemic_venous", 
    "EC_aerocyte", "EC_lymphatic",
   "Fibrotic_Macrophage", "Monocyte_derived_Macrophage", "Alveolar_Macrophage", 
    "Monocyte_nonClassical", "Monocyte_Classical", "DC2", "DC1", "Mast", 
    "IFNresp_Monocyte_derived_Macrophage", "pDC", "Langerhans-Cells", 
    "Macrophage_Cycling", "Neutro",
   "CD4+_TC_naive", "CD4+_TC_effector_memory", "CD4+_TC_TH1", "Treg", 
    "NK_Cells_CD16+", "CD8+_TC_resident_memory", "B_Cells_memory", 
    "CD8+_TC_effector_memory", "ILC3", "gamma_delta_TC", "CD4+_TC_TH17", 
    "CD4+_TC_Cytotoxic", "B_Cells_naive", "Plasma_Cells", "NK_Cells_CD16-")
)
```
### Standardize Cohort Annotation for Heatmap
```{r}
test$orig.ident_final %>% unique()

Batch2_Fren_IDs <- c("PPFE_PPFE21", 
 "PPFE_PPFE022", 
 "PPFE_PPFE025", 
 "PPFE_PPFE026", 
 "PPFE_PPFE027", 
 "PPFE_PPFE028", 
 "PPFE_PPFE029", 
 "PPFE_PPFE030", 
 "PPFE_PPFE031", 
 "PPFE_PPFE032", 
 "PPFE_PPFE033", 
 "PPFE_PPFE034", 
 "PPFE_PPFE035", 
 "PPFE_PPFE037", 
 "PPFE_PPFE038", 
 "PPFE_PPFE040", 
 "PPFE_PPFE023", 
 "PPFE_PPFE024", 
 "PPFE_PPFE036", 
 "PPFE_PPFE039")

BAtch1_Han_IDs <- c("PPFE_PPFE001" , "PPFE_PPFE002" ,  "PPFE_PPFE003" ,  "PPFE_PPFE004" ,  "PPFE_PPFE005" , "PPFE_PPFE006"  , "PPFE_PPFE007"  ,   "PPFE_PPFE008",   "PPFE_PPFE009",   "PPFE_PPFE010",   "PPFE_PPFE011",   "PPFE_PPFE012",   "PPFE_PPFE013",   "PPFE_PPFE014",   "PPFE_PPFE015a",  "PPFE_PPFE015b",  "PPFE_PPFE016",   "PPFE_PPFE017",   "PPFE_PPFE018",   "PPFE_PPFE019",   "PPFE_PPFE020", 
                    "PPFE_PPFE40", 
                    "PPFE_PPFE41", 
                    "PPFE_PPFE42",
                    "PPFE_PPFE43",
                    "PPFE_PPFE44")

CTRL_ids <- c("CTRL_GLE822" ,  "CTRL_GLE891"   ,"CTRL_CTRL002",  "CTRL_GLR632"   ,"CTRL_GLR635",   "CTRL_CTRL007",  "CTRL_GLE639"  , "CTRL_GLR609" ,  "CTRL_GLR725"  ,"CTRL_CTR048"  , "CTRL_CTR068" ,  "CTRL_CTR069" ,  "CTRL_CTR070", "CTRL_CTR058", "CTRL_CTR071" , "CTRL_CTR051")



test <- test %>%
 mutate(Cohort = case_when(orig.ident_final %in% Batch2_Fren_IDs ~ "France",
                            orig.ident_final %in% BAtch1_Han_IDs ~ "Germany",
                            orig.ident_final %in% CTRL_ids ~ "Germany",
                            .default = "Belgium")
        )




test <- test %>%
  mutate(Cohort = case_when(orig.ident_final == "PPFE040" ~ "Germany",
                            .default = Cohort))
table(test$orig.ident_final, test$Cohort)

```
### Start Heatmap-Script
```{r}
### with multithreading, first make all possible combinations
FullDataset_Object <-  test  #!= "KroBan")


celltypes_to_plot <- levels(as.factor(test$Celltype_ID_final_heat))

FullDataset_Object$cellBarcode <- colnames(FullDataset_Object)

cellTypes <- levels(as.factor(FullDataset_Object$Celltype_ID_final_heat))

cellTypes <- cellTypes[cellTypes %in% celltypes_to_plot]

Lineages <- levels(test$Lineage)

meta.data.sub <- FullDataset_Object@meta.data[,c("Cohort",
                                                 "disease.ident", # cohort.ident condition
                                                 "Celltype_ID_final_heat", # cell.type.ident
                                                 "orig.ident_final", # subject.ident
                                                 "cellBarcode")]



get.CT.DS.subj.vector <- function(cellTypes){
  
  tmp.meta.data <- meta.data.sub %>% dplyr::filter(Celltype_ID_final_heat == cellTypes)
  
  cohorts <- unique(tmp.meta.data$Cohort)
  
  disease <- unique(tmp.meta.data$disease.ident)
  
  subjects <- unique(tmp.meta.data$orig.ident_final)
  
  tmp.CT.DS.subj.vector <- vector()
  
  for(j in 1:length(cohorts)){
    
    for(k in 1:length(disease)){
      
      for(l in 1:length(subjects)){
      
      temp.cells <- tmp.meta.data %>% dplyr::filter(Cohort==cohorts[j] & disease.ident==disease[k] & orig.ident_final==subjects[l]) %>% pull(cellBarcode)
      
      if ( length(temp.cells) >15 ) { # >=
        
        tmp.CT.DS.subj.vector <- c(tmp.CT.DS.subj.vector, paste(cellTypes, cohorts[j], disease[k], subjects[l], sep="__"))
        
        }
      }
      
    }

  }
  
  cat("Completed for ", cellTypes, ".\n", sep="")
  
  return(tmp.CT.DS.subj.vector)
  
}

celltype_cohort_subject.list <- parallel::mclapply(cellTypes, get.CT.DS.subj.vector, mc.cores=1)

celltype_cohort_subject <- unlist(celltype_cohort_subject.list)
celltype_cohort_subject %>% tail(300)
```
```{r}
library(Matrix)

DefaultAssay(FullDataset_Object) <- "RNA"

get.SubjectcohortCellTypeAvg <- function(celltype_cohort_subject){
  
  temp.cell.type <- strsplit(as.character(celltype_cohort_subject),"__")[[1]][1]
  
  temp.cohort <- strsplit(as.character(celltype_cohort_subject),"__")[[1]][2]
  
  temp.disease <- strsplit(as.character(celltype_cohort_subject),"__")[[1]][3]
  
  temp.subject <- strsplit(as.character(celltype_cohort_subject),"__")[[1]][4]
  
  temp.meta.data <- FullDataset_Object@meta.data[,c("Cohort", "disease.ident", "Celltype_ID_final_heat", "orig.ident_final", "cellBarcode")]
  
  temp.cells <- temp.meta.data %>% dplyr::filter(Celltype_ID_final_heat==temp.cell.type & Cohort==temp.cohort & disease.ident==temp.disease &
                                            
                                            orig.ident_final==temp.subject) %>% pull(cellBarcode)
  
  if (length(temp.cells) > 15) {
    
    tmp.df <- as.data.frame(rowMeans(GetAssayData(FullDataset_Object)[,temp.cells]))       
    
  } else {
    
    tmp.df <- as.data.frame(GetAssayData(FullDataset_Object)[,temp.cells])
    
    cat("Subject",temp.subject,"only has 1",temp.cell.type,"cell, using singlet for",temp.cohort,"representation...\n",sep=" ")
    
  }
  
  colnames(tmp.df) <- paste(celltype_cohort_subject)
  
  return(tmp.df)
  
}



collapsed.mtx.list <- parallel::mclapply(celltype_cohort_subject, get.SubjectcohortCellTypeAvg, mc.cores=1)

collapsed.SubjectcohortCellTypeAvg.mtx <- Matrix(as.matrix(do.call(cbind, collapsed.mtx.list)), sparse = TRUE)

dim(collapsed.SubjectcohortCellTypeAvg.mtx) # Average for each subject for each gene 

```
```{r}
heatmap_metadata <- as.data.frame(cbind(colnames(collapsed.SubjectcohortCellTypeAvg.mtx),
                                        
                                        sapply(strsplit(as.character(colnames(collapsed.SubjectcohortCellTypeAvg.mtx)),"__"), `[`, 1),
                                        
                                        sapply(strsplit(as.character(colnames(collapsed.SubjectcohortCellTypeAvg.mtx)),"__"), `[`, 2),
                                        
                                        sapply(strsplit(as.character(colnames(collapsed.SubjectcohortCellTypeAvg.mtx)),"__"), `[`, 3),
                                        
                                        sapply(strsplit(as.character(colnames(collapsed.SubjectcohortCellTypeAvg.mtx)),"__"), `[`, 4)
                                        )
                                  )


colnames(heatmap_metadata) <- c("cell.ident","cell.type.ident", "cohort.ident", "disease.ident", "subject.ident")


heatmap_metadata
```
```{r}
heatmap_metadata$cell.type.ident <- factor(heatmap_metadata$cell.type.ident, levels=celltypes_to_plot)

heatmap_metadata$cohort.ident <- factor(heatmap_metadata$cohort.ident, levels=c("Germany", "France","Belgium"))

heatmap_metadata$disease.ident <- factor(heatmap_metadata$disease.ident, levels=c( "CTRL", "PPFE", "IPF"))

heatmap_metadata
```
```{r}
cell_order <- heatmap_metadata %>% arrange(cell.type.ident, cohort.ident, disease.ident, subject.ident ) %>% pull(cell.ident)

cohort_order <- heatmap_metadata %>% arrange(cell.type.ident, cohort.ident,  disease.ident, subject.ident ) %>% pull(cohort.ident)

disease_order <- heatmap_metadata %>% arrange(cell.type.ident, cohort.ident, disease.ident, subject.ident ) %>% pull(disease.ident)

celltype_order <- heatmap_metadata %>% arrange(cell.type.ident, cohort.ident,  disease.ident, subject.ident ) %>% pull(cell.type.ident)

subject_order <- heatmap_metadata %>% arrange(cell.type.ident, cohort.ident,  disease.ident, subject.ident ) %>% pull(subject.ident )

celltype_order %>% levels()
```

```{r}
heatmap_df1 <-  as.matrix(collapsed.SubjectcohortCellTypeAvg.mtx[Lineage_genes,cell_order])
heatmap_df <- heatmap_df1
```

### Normalize by Cohort: Hannover vs. Leuven
```{r}
ppfe_cells <- heatmap_metadata %>% 
    dplyr::filter(cohort.ident %in% c("Germany", "France")) %>% 
    pull(cell.ident)

ipf_cells <- heatmap_metadata %>% dplyr::filter(cohort.ident == "Belgium") %>% pull(cell.ident)

heatmap_df_ppfe <- heatmap_df[, ppfe_cells]
heatmap_df_ipf <- heatmap_df[, ipf_cells]


heatmap_metadata %>% dplyr::filter(cohort.ident == "Belgium")
```




```{r}
myUnityNormalize <- function(x){(x-min(x))/(max(x)-min(x))}

# Normalize each subset
heatmap_df_ppfe_normalized <- t(apply(heatmap_df_ppfe, MARGIN = 1, FUN = myUnityNormalize))
heatmap_df_ipf_normalized <- t(apply(heatmap_df_ipf, MARGIN = 1, FUN = myUnityNormalize))


# Combine both normalized datasets
heatmap_df_normalized <- cbind(heatmap_df_ppfe_normalized, heatmap_df_ipf_normalized)

heatmap_df_normalized <- heatmap_df_normalized[,cell_order]  # Originale Reihenfolge wiederherstellen
```


### Create Ordering for column split
```{r, fig.width=20, fig.height=9.5}
Epithelial = c(
    "AEC2", "AEC1", "Club", "AEC_intermediate", "Basal", "Secretory", 
    "Aberrant_Basaloid", "Goblet", "Ciliated", "PNEC", "Mesothelium","Schwann_Neu"
  )
Mesenchymal = c(
    "Pericytes", "Alveolar_Fb", "Adventitial_Fb", "Subpleural_Fb", 
    "Adventitial_like_Fb", "Smooth_Muscle", "CTHRC1_MyoFb", "Airway_Fb"
  )
Endothelial = c(
    "EC_pulmonary_venous", "EC_arterial", "EC_general_cap", "EC_systemic_venous", 
    "EC_aerocyte", "EC_lymphatic"
  )
Myeloids = c(
    "Fibrotic_Macrophage", "Monocyte_derived_Macrophage", "Alveolar_Macrophage", 
    "Monocyte_nonClassical", "Monocyte_Classical", "DC2", "DC1", "Mast", 
    "IFNresp_Monocyte_derived_Macrophage", "pDC", "Langerhans-Cells", 
    "Macrophage_Cycling", "Neutro"
  )
Lymphocytes = c(
    "CD4+_TC_naive", "CD4+_TC_effector_memory", "CD4+_TC_TH1", "Treg", 
    "NK_Cells_CD16+", "CD8+_TC_resident_memory", "B_Cells_memory", 
    "CD8+_TC_effector_memory", "ILC3", "gamma_delta_TC", "CD4+_TC_TH17", 
    "CD4+_TC_Cytotoxic", "B_Cells_naive", "Plasma_Cells", "NK_Cells_CD16-"
  )

split_order <- celltype_order 

# Function to rename based on occurrence in groups
rename_celltypes <- function(celltypes, Epithelial, Mesenchymal, Endothelial, Myeloids, Lymphocytes) {
  sapply(celltypes, function(ct) {
    if (ct %in% Epithelial) {
      return("Epithelium")
    } else if (ct %in% Mesenchymal) {
      return("Mesenchyme")
    } else if (ct %in% Endothelial) {
      return("Endothelium")
    } else if (ct %in% Myeloids) {
      return("Myeloids")
    } else if (ct %in% Lymphocytes) {
      return("Lymphocytes")
    } else {
      return("Unknown")
    }
  })
}

# Rename cell types
renamed_celltypes <- rename_celltypes(celltype_order, Epithelial, Mesenchymal, Endothelial, Myeloids, Lymphocytes)

# View renamed cell types
renamed_celltypes %>% unique()

renamed_celltypes <- factor(renamed_celltypes, levels = c("Epithelium", "Mesenchyme", "Endothelium", "Myeloids", "Lymphocytes"))

# Check final ordering of factor levels
renamed_celltypes %>% levels()
```


```{r}
library(ComplexHeatmap)
#dis_pal <- list("royalblue","red3", "darkgreen")
#cohort_pal <- list("darkgrey", "lightgrey")
cohort_pal_heat <- palette_scp(x = test$Cohort, palcolor = c("grey88", "grey58", "grey18"))

cohort_colors <- palette_scp(x = cohort_order, palcolor = cohort_pal_heat, n = 3, matched = F) # disease.pal
disease_colors <- palette_scp(x = disease_order, palcolor = pal_dis_pal, n = 3, matched = F) # disease.pal
celltype_colors <- palette_scp(x = celltype_order, palcolor = pal_cluster_scp_pal, matched = F)
subject_colors <- SCP::palette_scp(x = subject_order,  palcolor = pal_subject_pal, n = 111, matched = F)
lineage_colors <- SCP::palette_scp(x = renamed_celltypes, palcolor = lin_pal, n = 5, matched = F)


# Plot Annotation without with legend
heatmap_cohort_annotation <- HeatmapAnnotation(lineage = renamed_celltypes, cell_type=celltype_order, cohort=cohort_order, disease = disease_order, subject=subject_order, col = list(cohort=cohort_colors, disease = disease_colors, lineage = lineage_colors, cell_type=celltype_colors, subject=subject_colors), show_legend = c("subject" = FALSE))

# Plot Annotation without any legend
heatmap_cohort_annotation <- HeatmapAnnotation(lineage = renamed_celltypes, 
                                               cell_type=celltype_order,
                                               cohort=cohort_order, 
                                               disease = disease_order, 
                                               subject=subject_order, 
                                               col = list(cohort=cohort_colors, 
                                                          disease = disease_colors, 
                                                          lineage = lineage_colors, 
                                                          cell_type=celltype_colors, 
                                                          subject=subject_colors),
                                               show_legend = F)#c("subject" = FALSE))

```

```{r, fig.width=20, fig.height=9.5}
library(viridis)
p <- Heatmap(heatmap_df_normalized, name = "Scaled Expression",
        col = viridis(256), 
        cluster_rows = F, 
        show_heatmap_legend = F,
        cluster_columns = F, 
        show_column_names = FALSE, 
        top_annotation=heatmap_cohort_annotation, 
        column_split=renamed_celltypes, 
        #row_split = c(rep("1. Celltype Marker", nrow(heatmap_df1)), rep("2. Aberrant Basaloid Marker", nrow(heatmap_df2))),  # split to create visual separation
       # row_gap = unit(c(2, 2), "mm"),  # larger gap around the separation line
        column_title = NULL, 
        use_raster=FALSE,
        column_gap = unit(0.5, "mm"))
p
```



```{r, fig.width=20, fig.height=9.5}
png(filename = file.path(folder, "PPFE_IPF_ALL_LINEAGES_HEATMAP_RPCA_Aberrant_Features_bycohort_recol.png"), height = 5, width = 15, units = "in", res = 1200)
#pdf("PPFE_MES_UMAP.pdf", height = 7, width = 7)
p
dev.off()
p

png(filename = file.path(folder, "PPFE_IPF_ALL_LINEAGES_HEATMAP_RPCA_Aberrant_Features_bycohort_narrow.png"), height = 5, width = 9, units = "in", res = 1200)
#pdf("PPFE_MES_UMAP.pdf", height = 7, width = 7)
p
dev.off()
p

pdf(file.path(folder, "PPFE_IPF_ALL_LINEAGES_HEATMAP_RPCA_Aberrant_Features_bycohort_recol.pdf") ,height = 7, width = 3.0)
p
dev.off()


pdf(file = file.path(folder, "PPFE_IPF_ALL_LINEAGES_HEATMAP_RPCA_Aberrant_Features_bycohort_legend.pdf"), height = 9.5, width = 20)
#pdf("PPFE_MES_UMAP.pdf", height = 7, width = 7)
p
dev.off()

```



# Save Final Dataset
```{r}
saveRDS(test, "PPFE_Germany_France_CTRL_IPF.int.11032025.rds")
```

























#========
## UnityHeatmap


### Celltype_ID_clean_coarse
```{r}


genes_for_heatmap <- c( 
    "ADRA1A", "LIMCH1", "MME", "FRAS1", "COL13A1", # "Alveolar_Fb"  
    "COL10A1", "POSTN", "CTHRC1", "FAP", "INHBA",  # "MyoFb_CTHRC1"
    "TSPAN11", "DIO2", "COL15A1", "LGR5", "TMEM132C", # Adv-like Fb
    "MFAP5", "SFRP1", "SFRP2", "SCARA5", "IGFBP3", "PI16", "VIT",  # ADventitial_Fb
    "MEDAG", "ZNF331","PDPN", "WT1", "HAS1", "PTX3", #  "Subpleural_Fb_HAS1"
    "DACH2", "WNT5A", "ITGA10", "NRCAM", "ASPN", # Airway_Fb
    "SPEG", "DES",  "MYH11", "LDB3", "ACTG2", #"Smooth_Muscle" 
    "CACNA1H","LAMC3", "HIGD1B", "COX4I2", "KCNK3"#,  # "Pericytes"  
    #"CALB2", "ITLN1", "MSLN", "CLDN1", "HP", # "Mesothelium" 
    #"L1CAM", "SOX10","GFRA3", "HAND2", "NRXN1" # Schwann
    )
```



```{r}
### with multithreading, first make all possible combinations

FullDataset_Object <- PPFE_endo
FullDataset_Object <- FullDataset_Object %>%
  mutate(Cohort = case_when(Project == "PPFE" ~ Cohort,
                                  Project == "IPF" ~ "Belgium")
       )
FullDataset_Object$Celltype_ID_coarse <- factor(FullDataset_Object$Celltype_ID_coarse, levels = c("EC_arterial", "EC_aerocyte", "EC_general_cap", "EC_pulmonary_venous", "EC_systemic_venous", "EC_lymphatic"))

celltypes_to_plot <- levels(FullDataset_Object$Celltype_ID_coarse)

FullDataset_Object$cellBarcode <- colnames(FullDataset_Object)

cellTypes <- levels(as.factor(FullDataset_Object$Celltype_ID_coarse))

cellTypes <- cellTypes[cellTypes %in% celltypes_to_plot]

meta.data.sub <- FullDataset_Object@meta.data[,c("Cohort",
                                                 "disease.ident", # cohort.ident condition
                                                 "Celltype_ID_coarse", # cell.type.ident
                                                 "orig.ident", # subject.ident
                                                 "cellBarcode")]

celltypes_to_plot

meta.data.sub$disease.ident %>% levels()
```
```{r}
### with multithreading, first make all possible combinations
FullDataset_Object <- PPFE_mes
FullDataset_Object <- FullDataset_Object %>%
  mutate(Cohort = case_when(Project == "PPFE" ~ Cohort,
                                  Project == "IPF" ~ "Belgium")
       )
FullDataset_Object$Celltype_ID_coarse <- factor(FullDataset_Object$Celltype_ID_coarse, levels = c("Alveolar_Fb", 
                                                                                                  "CTHRC1_MyoFb", 
                                                                                                  "Adventitial_like_Fb",
                                                                                                  "Adventitial_Fb", 
                                                                                                  "Subpleural_Fb", 
                                                                                                  "Airway_Fb", 
                                                                                                  "Smooth_Muscle", 
                                                                                                  "Pericytes", 
                                                                                                  "Mesothelium", 
                                                                                                  "Schwann_Neu"))

celltypes_to_plot <- levels(FullDataset_Object$Celltype_ID_coarse)

FullDataset_Object$cellBarcode <- colnames(FullDataset_Object)

cellTypes <- levels(as.factor(FullDataset_Object$Celltype_ID_coarse))

cellTypes <- cellTypes[cellTypes %in% celltypes_to_plot]

meta.data.sub <- FullDataset_Object@meta.data[,c("Cohort",
                                                 "disease.ident", # cohort.ident condition
                                                 "Celltype_ID_coarse", # cell.type.ident
                                                 "orig.ident", # subject.ident
                                                 "cellBarcode")]



get.CT.DS.subj.vector <- function(cellTypes){
  
  tmp.meta.data <- meta.data.sub %>% dplyr::filter(Celltype_ID_coarse == cellTypes)
  
  cohorts <- unique(tmp.meta.data$Cohort)
  
  disease <- unique(tmp.meta.data$disease.ident)
  
  subjects <- unique(tmp.meta.data$orig.ident)
  
  tmp.CT.DS.subj.vector <- vector()
  
  for(j in 1:length(cohorts)){
    
    for(k in 1:length(disease)){
      
      for(l in 1:length(subjects)){
      
      temp.cells <- tmp.meta.data %>% dplyr::filter(Cohort==cohorts[j] & disease.ident==disease[k] & orig.ident==subjects[l]) %>% pull(cellBarcode)
      
      if ( length(temp.cells) >15 ) { # >=
        
        tmp.CT.DS.subj.vector <- c(tmp.CT.DS.subj.vector, paste(cellTypes, cohorts[j], disease[k], subjects[l], sep="__"))
        
        }
      }
      
    }

  }
  
  cat("Completed for ", cellTypes, ".\n", sep="")
  
  return(tmp.CT.DS.subj.vector)
  
}

celltype_cohort_subject.list <- parallel::mclapply(cellTypes, get.CT.DS.subj.vector, mc.cores=1)

celltype_cohort_subject <- unlist(celltype_cohort_subject.list)
celltype_cohort_subject
```
```{r}
library(Matrix)

DefaultAssay(FullDataset_Object) <- "RNA"

get.SubjectcohortCellTypeAvg <- function(celltype_cohort_subject){
  
  temp.cell.type <- strsplit(as.character(celltype_cohort_subject),"__")[[1]][1]
  
  temp.cohort <- strsplit(as.character(celltype_cohort_subject),"__")[[1]][2]
  
  temp.disease <- strsplit(as.character(celltype_cohort_subject),"__")[[1]][3]
  
  temp.subject <- strsplit(as.character(celltype_cohort_subject),"__")[[1]][4]
  
  temp.meta.data <- FullDataset_Object@meta.data[,c("Cohort", "disease.ident", "Celltype_ID_coarse", "orig.ident", "cellBarcode")]
  
  temp.cells <- temp.meta.data %>% dplyr::filter(Celltype_ID_coarse==temp.cell.type & Cohort==temp.cohort & disease.ident==temp.disease &
                                            
                                            orig.ident==temp.subject) %>% pull(cellBarcode)
  
  if (length(temp.cells) > 15) {
    
    tmp.df <- as.data.frame(rowMeans(GetAssayData(FullDataset_Object)[,temp.cells]))       
    
  } else {
    
    tmp.df <- as.data.frame(GetAssayData(FullDataset_Object)[,temp.cells])
    
    cat("Subject",temp.subject,"only has 1",temp.cell.type,"cell, using singlet for",temp.cohort,"representation...\n",sep=" ")
    
  }
  
  colnames(tmp.df) <- paste(celltype_cohort_subject)
  
  return(tmp.df)
  
}



collapsed.mtx.list <- parallel::mclapply(celltype_cohort_subject, get.SubjectcohortCellTypeAvg, mc.cores=1)

collapsed.SubjectcohortCellTypeAvg.mtx <- Matrix(as.matrix(do.call(cbind, collapsed.mtx.list)), sparse = TRUE)

dim(collapsed.SubjectcohortCellTypeAvg.mtx) # Average for each subject for each gene 

```
```{r}
heatmap_metadata <- as.data.frame(cbind(colnames(collapsed.SubjectcohortCellTypeAvg.mtx),
                                        
                                        sapply(strsplit(as.character(colnames(collapsed.SubjectcohortCellTypeAvg.mtx)),"__"), `[`, 1),
                                        
                                        sapply(strsplit(as.character(colnames(collapsed.SubjectcohortCellTypeAvg.mtx)),"__"), `[`, 2),
                                        
                                        sapply(strsplit(as.character(colnames(collapsed.SubjectcohortCellTypeAvg.mtx)),"__"), `[`, 3),
                                        
                                        sapply(strsplit(as.character(colnames(collapsed.SubjectcohortCellTypeAvg.mtx)),"__"), `[`, 4)
                                        )
                                  )


colnames(heatmap_metadata) <- c("cell.ident","cell.type.ident", "cohort.ident", "disease.ident", "subject.ident")


heatmap_metadata$disease.ident %>% unique()
```
```{r}
heatmap_metadata$cell.type.ident <- factor(heatmap_metadata$cell.type.ident, levels=celltypes_to_plot)

heatmap_metadata$cohort.ident <- factor(heatmap_metadata$cohort.ident, levels=c("Germany", "France","Belgium"))

heatmap_metadata$disease.ident <- factor(heatmap_metadata$disease.ident, levels=c( "CTRL", "PPFE", "IPF"))

heatmap_metadata
```
```{r}
cell_order <- heatmap_metadata %>% arrange(cell.type.ident, cohort.ident, disease.ident, subject.ident ) %>% pull(cell.ident)

cohort_order <- heatmap_metadata %>% arrange(cell.type.ident, cohort.ident,  disease.ident, subject.ident ) %>% pull(cohort.ident)

disease_order <- heatmap_metadata %>% arrange(cell.type.ident, cohort.ident, disease.ident, subject.ident ) %>% pull(disease.ident)

celltype_order <- heatmap_metadata %>% arrange(cell.type.ident, cohort.ident,  disease.ident, subject.ident ) %>% pull(cell.type.ident)

subject_order <- heatmap_metadata %>% arrange(cell.type.ident, cohort.ident,  disease.ident, subject.ident ) %>% pull(subject.ident )
```




```{r}
myUnityNormalize <- function(x){(x-min(x))/(max(x)-min(x))}


ppfe_cells <- heatmap_metadata %>% 
    dplyr::filter(cohort.ident %in% c("Germany", "France")) %>% 
    pull(cell.ident)
ipf_cells <- heatmap_metadata %>% dplyr::filter(cohort.ident == "Belgium") %>% pull(cell.ident)

heatmap_df1 <-  as.matrix(collapsed.SubjectcohortCellTypeAvg.mtx[genes_for_heatmap,cell_order])
heatmap_df <- heatmap_df1

heatmap_df_ppfe <- heatmap_df[, ppfe_cells]
heatmap_df_ipf <- heatmap_df[, ipf_cells]

heatmap_df_ppfe_normalized <- t(apply(heatmap_df_ppfe, MARGIN = 1, FUN = myUnityNormalize))
heatmap_df_ipf_normalized <- t(apply(heatmap_df_ipf, MARGIN = 1, FUN = myUnityNormalize))

# Combine both normalized datasets
heatmap_df_normalized <- cbind(heatmap_df_ppfe_normalized, heatmap_df_ipf_normalized)
heatmap_df_normalized <- heatmap_df_normalized[genes_for_heatmap,cell_order]  # Originale Reihenfolge wiederherstellen


library(ComplexHeatmap)
dis_pal <- list("royalblue","red3", "darkgreen")
#cohort_pal <- list("darkgrey", "lightgrey")
cohort_colors <- palette_scp(x = cohort_order, palcolor = pal_cohort_pal_heat, n = 3, matched = F) # disease.pal
disease_colors <- palette_scp(x = disease_order, palcolor = pal_dis_pal, n = 3, matched = F) # disease.pal
#celltype_colors <- palette_scp(x = celltype_order, palette = "Paired", n = 6, matched = F)
subject_colors <- SCP::palette_scp(x = subject_order,  palcolor = pal_subject_pal, n = 64, matched = F)

# Use standardized palettes among project1
celltype_colors <- pal_cluster_scp_pal
#subject_colors <-  pal_subject_pal
#disease_colors <-  pal_dis_pal
#cohort_colors <-   pal_cohort_pal

# Plot
heatmap_cohort_annotation <- HeatmapAnnotation(cell_type=celltype_order, cohort=cohort_order, disease = disease_order, subject=subject_order,
                                               
                                               col = list(cohort=cohort_colors, disease = disease_colors, cell_type=celltype_colors, subject=subject_colors),
                                               
                                               show_legend = c("subject" = FALSE), 
                                               annotation_label = c("Celltype", "Cohort", "Disease", "Subject")
)



```

```{r, fig.width=12, fig.height=9.5}
library(viridis)
p <- Heatmap(heatmap_df_normalized, name = "Scaled_Expression",
        col = inferno(256), 
        cluster_rows = F, 
        cluster_columns = F, 
        show_column_names = FALSE, 
        top_annotation=heatmap_cohort_annotation, 
        column_split=celltype_order, 
       # row_split = c(rep("1. Celltype Marker", nrow(heatmap_df1)), rep("2. Aberrant Basaloid Marker", nrow(heatmap_df2))),  # split to create visual separation
      #  row_gap = unit(c(2, 2), "mm"),  # larger gap around the separation line
        column_title = NULL, 
        use_raster=FALSE,
        column_gap = unit(0.25, "mm"))

p
```
```{r}
heatmap_df1 <-  as.matrix(collapsed.SubjectcohortCellTypeAvg.mtx[genes_for_heatmap,cell_order])

lineage_for_heatmap <- c("ELN", #"LOX", "EMILIN1", 
                            "FBN1", #"FBN2", 
                            "FBLN1",
                            "FBLN2", #"FBLN5", 
                            "LOXL1", "LOXL2", #"LOXL3", "LOXL4", 
                            #"MFAP1",
                            "MFAP2",#"MFAP3",
                            "MFAP4","MFAP5")

heatmap_df2 <- as.matrix(collapsed.SubjectcohortCellTypeAvg.mtx[lineage_for_heatmap,cell_order])

separation_row <- matrix(NA, nrow = 1, ncol = ncol(heatmap_df2))

#heatmap_df <- heatmap_df1
heatmap_df <- rbind(heatmap_df2, heatmap_df1)
```





```{r}
ppfe_cells <- heatmap_metadata %>% 
    dplyr::filter(cohort.ident %in% c("Germany", "France")) %>% 
    pull(cell.ident)
ipf_cells <- heatmap_metadata %>% dplyr::filter(cohort.ident == "Belgium") %>% pull(cell.ident)

heatmap_df_ppfe <- heatmap_df[, ppfe_cells]
heatmap_df_ipf <- heatmap_df[, ipf_cells]

```




```{r}
myUnityNormalize <- function(x){(x-min(x))/(max(x)-min(x))}

# Normalize each subset
heatmap_df_ppfe_normalized <- t(apply(heatmap_df_ppfe, MARGIN = 1, FUN = myUnityNormalize))
heatmap_df_ipf_normalized <- t(apply(heatmap_df_ipf, MARGIN = 1, FUN = myUnityNormalize))


# Combine both normalized datasets
heatmap_df_normalized <- cbind(heatmap_df_ppfe_normalized, heatmap_df_ipf_normalized)

heatmap_df_normalized <- heatmap_df_normalized[,cell_order]  # Originale Reihenfolge wiederherstellen
```




```{r}
#library(ComplexHeatmap)
dis_pal <- list("royalblue","red3", "darkgreen")
#cohort_pal <- list("darkgrey", "lightgrey")
cohort_colors <- palette_scp(x = cohort_order, palcolor = pal_cohort_pal_heat, n = 3, matched = F) # disease.pal
disease_colors <- palette_scp(x = disease_order, palcolor = pal_dis_pal, n = 3, matched = F) # disease.pal
#celltype_colors <- palette_scp(x = celltype_order, palette = "Paired", n = 6, matched = F)
subject_colors <- SCP::palette_scp(x = subject_order,  palcolor = pal_subject_pal, n = 64, matched = F)

# Use standardized palettes among project1
celltype_colors <- pal_cluster_scp_pal
#subject_colors <-  pal_subject_pal
#disease_colors <-  pal_dis_pal
#cohort_colors <-   pal_cohort_pal

heatmap_cohort_annotation <- HeatmapAnnotation(cell_type=celltype_order, cohort=cohort_order, disease = disease_order, subject=subject_order,
                                               
                                               col = list(cohort=cohort_colors, disease = disease_colors, cell_type=celltype_colors, subject=subject_colors),
                                               
                                               show_legend = c("subject" = FALSE))
```

```{r, fig.width=7, fig.height=9.5}
library(viridis)
p <- Heatmap(heatmap_df_normalized, name = "Normalized Expression" ,
        col = inferno(256), 
        cluster_rows = F, 
        cluster_columns = F, 
        show_column_names = FALSE, 
        top_annotation=heatmap_cohort_annotation, 
        column_split=celltype_order, 
        row_split = c(rep("1. Elastic Fibre Metabolism", nrow(heatmap_df2)), rep("2. Celltype Marker", nrow(heatmap_df1))),  # split to create visual separation
        row_gap = unit(c(2, 2), "mm"),  # larger gap around the separation line
        column_title = NULL, 
        use_raster=FALSE,
        column_gap = unit(0.25, "mm"))

p
```
```{r}
png(file.path(folder, "Figure_4_Heatmap_Final.png"), units = "in", width = 14, height = 11, res = 1200)
p
dev.off()

pdf(file.path(folder, "Figure_4_Heatmap_Final.pdf"), width = 14, height = 11)
p
dev.off()
```
