---
title: "Figure_1_Plots"
author: "Jannik Ruwisch"
date: "2025-01-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      fig.width = 14,
                      fig.height = 7,
                      #dpi = 1200,
                      dev = "png",
                      cache = F)
```

# Load the Packages
```{r}
library(Seurat)
library(SeuratObject)
library(dplyr)
library(tidyverse)
library(tidyseurat)
library(ggpubr)
library(SCP)
```

# Read in the Object
```{r}
test <- readRDS("PPFE_cleaned_datasetes_ALL_RPCA_clean_integrated_22_12_2024.rds")
```

# Select Folder Path
```{r}
folder <- choose.dir()
```

# Define Color Palettes`
```{r}
pal_cluster_scp_pal <- c(
  # Epithelial Cells
  "#7F3C8D",  # AEC2
  "#199784",  # AEC1
  "#715588",  # Club
  "#8B8B60",  # AEC_intermediate
  "#A5AA99",  # Basal
  "#CC831E",  # Secretory
  "#ff084a",  # Aberrant_Basaloid | corrected
  "#446293",  # Goblet
  "#8BAC5C",  # Ciliated
  "#FDBF6F",  # PNEC
  "#FDBF6F",  # Mesothelium
  "#000000",  # Schwann_Neu (Not specified, placeholder)

  # Fibroblasts & Stromal Cells
  "#33A02C",  # Pericytes
  "#A6CEE3",  # Alveolar_Fb
  "#6A3D9A",  # Adventitial_Fb
  "#FB9A99",  # Subpleural_Fb
  "#CAB2D6",  # Adventitial_like_Fb
  "#B2DF8A",  # Smooth_Muscle
  "#E31A1C",  # CTHRC1_MyoFb
  "#1F78B4",  # Airway_Fb

  # Endothelial Cells
  "#4DBBD5",  # EC_pulmonary_venous
  "#F39B7F",  # EC_arterial
  "#3C5488",  # EC_general_cap
  "#E64B35",  # EC_systemic_venous
  "#8491B4",  # EC_aerocyte
  "#00A087",  # EC_lymphatic

  # Myeloid Cells
  "#BA4769", # Fibrotic_Macrophage"
  "#BE777C",  # Monocyte_derived_Macrophage
  "#FF6F00",  # Alveolar_Macrophage
  "#BA7769",  # Monocyte_nonClassical
  "#6A7998",  # Monocyte_Classical
  "#1D5152",  # DC2
  "#94CABB",  # DC1
  "#CB1700",  # Mast
  "#B7BBC9",  # IFNresp_Monocyte_derived_Macrophage
  "#58644F",  # pDC
  "#3F4041",  # Langerhans-Cells
  "#67549A",  # Macrophage_Cycling
  "#217885",  # Neutro

  # Lymphoid Cells
  "#E6F5B7",  # CD4+_TC_naive
  "#D9D7C9",  # CD4+_TC_effector_memory
  "#D9B382",  # CD4+_TC_TH1 (Not explicitly named, assumed TH1 effector memory)
  "#8DD3C7",  # Treg
  "#DED7DA",  # NK_Cells_CD16+
  "#C7D98C",  # CD8+_TC_resident_memory
  "#FFED6F",  # B_Cells_memory
  "#D8C965",  # CD8+_TC_effector_memory
  "#88ADCC",  # ILC3
  "#F9CDE4",  # gamma_delta_TC
  "#D9B382",  # CD4+_TC_TH17 (Assumed from similar CD4 TC subsets)
  "#D9B382",  # CD4+_TC_Cytotoxic (Not explicitly defined, assumed)
  "#D6EBB2",  # B_Cells_naive
  "#C2ADC0",  # Plasma_Cells
  "#C69FC7" )  # NK_Cells_CD16-
  
# Define Palettes for Heatmap
  lin_pal <- palette_scp(x = test$Lineage,palcolor = c( "#cc79a7", "#009e73", "#5654e9", "#d55e00"  , "#e69f00"))
  pal_dis_pal <- palette_scp(x = test$disease.ident%>% unique(), palcolor = c("royalblue", "red3", "#ffd59b"), matched = F)
  pal_cohort_pal_heat <- palette_scp(x = c("Germany", "France",  "Belgium"), palcolor = c("grey88", "grey58", "grey18"), matched = F)
  pal_cohort_pal <-  c("royalblue", "red1", "darkred", "#ffd59b")
  pal_subject_pal <-  palette_scp(palette = "igv", x = test$orig.ident_final %>% unique(), matched = F) 
```


# Plots

### Overall Dataset
```{r, fig.width=14, fig.height=7}
# Proportion / cell number composition per cluster
ggData = data.frame(prop.table(table(test$Lineage, test$orig.ident_final), margin = 2))
colnames(ggData) = c("Celltype_ID", "Orig.Ident",  "value")

ggData <- ggData %>%
  mutate(Disease = substr(Orig.Ident, 1,4))

subset(test, disCohort_FINAL == "France_PPFE")$orig.ident_final %>% unique()-> Batch2_Fren_IDs 
subset(test, disCohort_FINAL == "Germany_PPFE")$orig.ident_final %>% unique()-> BAtch1_Han_IDs 
subset(test, disCohort_FINAL == "Germany_CTRL")$orig.ident_final %>% unique()-> CTRL_ids
```

## Stacked Bar Plots
```{r, fig.width=14, fig.height=7}
ggData <- ggData %>%
 mutate(Disease = case_when(Orig.Ident %in% Batch2_Fren_IDs ~ "France_PPFE",
                            Orig.Ident %in% BAtch1_Han_IDs ~ "Germany_PPFE",
                            Orig.Ident %in% CTRL_ids ~ "Germany_CTRL",
                            .default = "Belgium_IPF")
        )

ggData$Disease <- factor(ggData$Disease, levels = c("Germany_CTRL", "France_PPFE", "Germany_PPFE", "Belgium_IPF"))

test$Celltype_ID_final_coarse %>% unique() -> Celltypes

patient_order <- ggData %>%
  dplyr::filter(Celltype_ID == 'Lymphoid') %>% 
  arrange(Disease, desc(value)) %>% 
  pull(Orig.Ident)

# Convert Patient_ID to a factor with the specific order
ggData$Orig.Ident <- factor(ggData$Orig.Ident, levels = patient_order)

# Calculate median values for each Celltype_ID within each Disease group
medians <- ggData %>%
  dplyr::filter(Disease =="Germany_PPFE") %>%
  group_by(Disease, Celltype_ID) %>%
  summarize(median_value = median(value)) %>%
  ungroup()

# Order Celltype_ID based on median values
ordered_celltypes <- medians %>%
  group_by(Celltype_ID) %>%
  summarize(overall_median = median(median_value)) %>%
  arrange(overall_median) %>%
  pull(Celltype_ID)
ordered_celltypes

# Update the levels of Celltype_ID in ggData
ggData$Celltype_ID <- factor(ggData$Celltype_ID, levels = c( "Epithelium","Mesenchyme","Endothelium", "Myeloid","Lymphoid")
                             )

# Plot with normalized bars
p <- ggplot(ggData, aes(x = Orig.Ident, y = value, fill = Celltype_ID)) +
  geom_bar(stat = "identity", position = "fill", size = 0.5, color = "black") +
  labs(title = "Stacked Bar Plot of Cell Type Frequencies by Patient (Ordered by Disease Cohort)",
       x = "Patient ID",
       y = "Proportion (%)") + 
  scale_fill_manual(values = lin_pal) +
  scale_color_manual(values = lin_pal) +
  theme_scp() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r, fig.width=30}
p <- ggplot(ggData, aes(x = Orig.Ident, y = value, fill = Celltype_ID)) +
  geom_bar(stat = "identity", position = "stack", size = 0.5, color = "black") +
  labs(#title = "Stacked Bar Plot of Cell Type Frequencies by Patient",
       x = NULL,
       y = "Frequency (%)") + 
  scale_fill_discrete(lin_pal) +
  scale_color_discrete(lin_pal) +
  scale_fill_manual(values = lin_pal) +
  scale_color_manual(values = lin_pal) +
  theme_scp() +
  theme(element_text(angle = 45, hjust = 1),
        axis.text.x.bottom = element_blank(), 
        axis.ticks.length.x = unit(0, "cm"))+
  facet_wrap(~Disease, scales = "free_x", ncol = 4) + NoLegend()+RotatedAxis()

# Export to PDF and PNG File
png((filename = file.path(folder,"Figure_1Stackedbarplots.png")), units = "in", width = 16, height = 6.5, res = 1200)
p
dev.off()

pdf((file = file.path(folder,"Figure_1Stackedbarplots.pdf")),width = 16, height = 6.5)
p
dev.off()
```

## Quantification of Remodeled/Healthy Parenchyma
```{r}
parquant <- readxl::read_xlsx("C:/Users/Jannik/OneDrive/PNE_MHH/AG_Prasse_OneDrive/PPFE/Automated_Image_Analysis/Figure_1_Quantification_of_Remodeled_Parenchyma/Alv_Remodeled_Par_Quantification_PPFE_Germany_France_FINAL_03052025.xlsx")

parquant$Orig.Ident <- parquant$ID

merged_df <- dplyr::left_join(ggData, parquant, by = "Orig.Ident")
merged_df <- merged_df %>%
  mutate(Remodeled_Alv_Ratio = ifelse(is.na(Remodeled_Alv_Ratio), 0, Remodeled_Alv_Ratio),
         Alv_Remodeled_Ratio = ifelse(is.na(Alv_Remodeled_Ratio), 0, Alv_Remodeled_Ratio),
         )

merged_df$Orig.Ident <- factor(merged_df$Orig.Ident, levels = patient_order)

# Normalize the top bars to be only 1/10 of the total height
merged_df <- merged_df %>%
  mutate(Remodeled_Alv_Scaled = Remodeled_Alv_Ratio / max(Remodeled_Alv_Ratio, na.rm = TRUE), 
         Alv_Remodeled_Scaled = Alv_Remodeled_Ratio / max(Alv_Remodeled_Ratio, na.rm = TRUE))

# Create the plot
p <- ggplot(merged_df, aes(x = Orig.Ident)) +
  
  # Main stacked bar plot (Cell Type Frequencies)
  geom_bar(aes(y = value, fill = Celltype_ID), stat = "identity", position = "fill", size = 0.5, color = "black") +

  # Add new y-scale for Remodeled data
  scale_y_continuous(
    name = "Cell Type Proportion (%)",
    sec.axis = sec_axis(~ . * max(merged_df$Remodeled_Alv_Ratio, na.rm = TRUE), name = "Remodeled Ratio (%)") # Secondary y-axis
  ) +
  
  # Secondary y-axis: Overlaying bars for Remodeled Alveoli and Alveolar Remodeling
  geom_col(aes(y = Remodeled_Alv_Scaled, fill = "Remodeled Alveoli"), alpha = 0.5, width = 0.5) +
  geom_col(aes(y = Alv_Remodeled_Scaled, fill = "Alveolar Remodeling"), alpha = 0.5, width = 0.5) +

  # Labels and themes
  labs(title = "Stacked Bar Plot with Remodeled Parenchyma",
       x = "Patient ID",
       y = "Proportion (%)") + 
  scale_fill_manual(values = c(lin_pal, 
                               "Remodeled Alveoli" = "darkviolet", 
                               "Alveolar Remodeling" = "orange")) +
  theme_scp() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y.right = element_text(color = "darkviolet"), # Highlight secondary axis
    axis.title.y.left = element_text(color = "black") # Keep main axis
  ) +
  facet_wrap(~Disease, scales = "free_x", ncol = 4) +
  NoLegend() + RotatedAxis()
```

```{r}
p1 <- ggplot(merged_df, aes(x = Orig.Ident, y = value*100, fill = Celltype_ID)) +
  geom_bar(stat = "identity", position = "stack", size = 0.5, color = "black") +
  labs(x = NULL,
       y = NULL) + 
  scale_fill_discrete(lin_pal) +
  scale_color_discrete(lin_pal) +
  scale_fill_manual(values = lin_pal) +
  scale_color_manual(values = lin_pal) +
  theme_scp() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~Disease, scales = "free_x", ncol = 4) + NoLegend()+RotatedAxis()

p1



plot_data <- merged_df %>%
  select(Orig.Ident,Disease,Celltype_ID, Alv_Remodeled_Ratio, Remodeled_Alv_Ratio) %>%
  pivot_longer(cols = c(Alv_Remodeled_Ratio, Remodeled_Alv_Ratio),
               names_to = "Classification",
               values_to = "Percentage") %>%
  filter(Celltype_ID == "Epithelium")

plot_data$Orig.Ident <- factor(plot_data$Orig.Ident, levels = patient_order)

# Create stacked bar plot
p2 <- ggplot(plot_data, aes(x = Orig.Ident, y = Percentage*100, fill = Classification)) +
  geom_bar(stat = "identity") +
  labs(#title = "Stacked Bar Plot of Percent Alv and Percent Remodeled per ID",
       x = NULL,
       y = NULL,
       fill = "Classification") +
  theme_scp() +
 # RotatedAxis()+
  scale_fill_manual(values = c("lightblue", "purple"))+
  scale_color_manual(values = c("lightblue", "purple"))+
  theme_scp() +
   theme(
    axis.text.x = element_blank(), 
    axis.text.x.bottom = element_blank(),  # Remove x-axis labels
    axis.title.x = element_blank(),  # Remove x-axis title
    axis.ticks.x = element_blank()#,  # Remove x-axis ticks
    #strip.text = element_blank() 
    )+ # Remove facet titles
  facet_wrap(~Disease, scales = "free_x", ncol = 4) + NoLegend()+RotatedAxis()+
  scale_y_continuous(breaks = c(0,100))
p2
```
```{r}
# Plot the Final Plots together with Stacked Bar Plots
pdf((file = file.path(folder,"Figure_1Stackedbarplots_lineage.pdf")), width = 17, height = 6.5)
p1
dev.off()

pdf((file = file.path(folder,"Figure_1Stackedbarplots_RemodAlv.pdf")), width = 17, height = 1.0)
p2
dev.off()
```

## Frequency Plots
```{r}
# Convert disease.ident and Celltype to factors
ggData$Disease <- as.factor(ggData$Disease)
ggData$Celltype_ID <- as.factor(ggData$Celltype_ID)

# Calculate the median frequency for each cell type
median_frequencies <- ggData %>%
  group_by(Celltype_ID) %>%
  summarize(median_freq = median(value)) %>%
  arrange(-median_freq)

# Reorder the Celltype factor levels based on median frequency
ggData$Celltype_ID <- factor(ggData$Celltype_ID, levels = median_frequencies$Celltype_ID)

# Plot
p <- ggplot(ggData, aes(x = Celltype_ID, y = value, fill = Disease)) +
  geom_boxplot() +
  labs(title = "Comparison of Cell Type Frequencies Between Disease Groups",
       x = "Cell Type",
       y = "Frequency (%)") +
  scale_fill_manual(values = pal_cohort_pal) +
  scale_color_manual(values = pal_cohort_pal) +
  theme_scp() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Add statistical comparisons
p <- p + stat_compare_means(aes(group = Disease), method = "t.test", label = "p.signif")

#==============================================

# Plot
p <- ggplot(ggData, aes(x = Celltype_ID, y = value, fill = Disease)) +
  geom_boxplot(outlier.shape = NA) +
  labs(title = "Relative Endothelial Celltype Frequencies",
       x = "Cell Type",
       y = "Frequency (%)") +
  scale_fill_manual(values = pal_cohort_pal) +
  scale_color_manual(values = pal_cohort_pal) +
  theme_scp() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Add statistical comparisons
p + stat_compare_means(aes(group = Disease), 
                       label = "p.signif")


png(filename = file.path(folder,"Figure_1_Boxplots.png"), height = 7, width = 9, units = "in", res = 1200)
p+ stat_compare_means(aes(group = Disease), 
                       label = "p.signif")
dev.off()

pdf(file.path(folder,"Figure_1_Boxplots.pdf"),  height = 7, width = 9)
p
dev.off()

#======================================
ggData_1 <- subset(ggData, Celltype_ID != "Alveolar_Fb" & 
                     Celltype_ID != "Smooth_Muscle" & 
                     Celltype_ID != "Adventitial_Fb" &
                     Celltype_ID != "CTHRC1_MyoFb" &
                     Celltype_ID != "Pericytes") 

# Calculate the median frequency for each cell type
median_frequencies <- ggData_1 %>%
  group_by(Celltype_ID) %>%
  summarize(median_freq = median(value)) %>%
  arrange(-median_freq)

# Reorder the Celltype factor levels based on median frequency
ggData_1$Celltype_ID <- factor(ggData_1$Celltype_ID, levels = median_frequencies$Celltype_ID)

# Plot
p <- ggplot(ggData_1, aes(x = Celltype_ID, y = value, fill = Disease)) +
  geom_boxplot(outlier.shape = NA
    ) +
  labs(title = NULL,
       x = "Cell Type",
       y = "Frequency (%)") +
  scale_fill_manual(values = pal_cohort_pal) +
  scale_color_manual(values =pal_cohort_pal) +
  theme_scp(aspect.ratio = 1) + 
  theme(axis.text.x = element_text(angle = 70, hjust = 1))#+ylim(0, 0.1)

# Add statistical comparisons
p1 <- p + stat_compare_means(aes(group = Disease),# method = "t.test", 
                            label = "p.signif")

# Display the plot
print(p1)

ggData_2 <- subset(ggData, Celltype_ID == "Alveolar_Fb" | 
                     Celltype_ID == "Smooth_Muscle" | 
                     Celltype_ID == "Adventitial_Fb" |
                     Celltype_ID == "CTHRC1_MyoFb" |
                     Celltype_ID == "Pericytes")

# Calculate the median frequency for each cell type
median_frequencies <- ggData_2 %>%
  group_by(Celltype_ID) %>%
  summarize(median_freq = median(value)) %>%
  arrange(-median_freq)

# Reorder the Celltype factor levels based on median frequency
ggData_2$Celltype_ID <- factor(ggData_2$Celltype_ID, levels = median_frequencies$Celltype_ID)

# Plot
p <- ggplot(ggData_2, aes(x = Celltype_ID, y = value, fill = Disease)) +
  geom_boxplot(outlier.shape = NA) +
  labs(title =NULL,
       x = "Cell Type",
       y = "Frequency (%)") +
 scale_fill_manual(values = pal_cohort_pal) +
  scale_color_manual(values = pal_cohort_pal) +
  theme_scp(aspect.ratio = 1) +
  theme(axis.text.x = element_text(angle = 70, hjust = 1))

# Add statistical comparisons
p2 <- p + stat_compare_means(aes(group = Disease), #method = "t.test", 
                            label = "p.signif")

# Display the plot
print(p2)

# Print together
p <- ggarrange(plotlist = list(p2, p1), ncol = 2, common.legend = T, align = "h")

pdf(file.path(folder,"Figure_6_Boxplots_split.pdf"),  height = 7, width = 9)
p
dev.off()
```

## UMAP - Celltype_ID_final
```{r}
p <- CellDimPlot(test, group.by = "Celltype_ID_final", raster = F, pt.size = 0.15, reduction = "ReUMAP_50", palette = "Paired", show_stat = F, xlab = "UMAP_1", ylab = "UAMP_2", legend.position = "none")+
  theme(axis.text = element_blank(),   # Remove axis text
        axis.ticks = element_blank())  # Remove axis ticks

pdf("PPFE_ALL_UMAP.pdf", height = 14, width = 14)
p
dev.off()
p
```

## UMAP - Lineage
```{r}
p <-DimPlot(test, group.by = c("Lineage"),
          shuffle = T, raster = F, label = F, combine = T, pt.size = 0.15,
          reduction = "ReUMAP_50", 
        cols =lin_pal)+ 
  theme_scp()+
  labs(x = "UMAP_1", y = "UMAP_2")+
  #NoLegend()+
  ggtitle("Subject")+
  scale_color_manual(values = as.vector(lin_pal))+
  theme(axis.text = element_blank(), legend.direction = "horizontal",   # Remove axis text
        axis.ticks = element_blank(),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 25, face = "bold"),
        legend.key.size = unit(2, "cm"),
        legend.position = "bottom")+
  guides(color = guide_legend(override.aes = list(size = 16)))  
        
         # Remove axis ticks

pdf(file.path(folder, "PPFE_ALL_MERGED_UMAP_Lineage_legend.pdf"), height = 14, width = 14)
p
dev.off()

```

## UMAP - Subject
```{r}
p <- DimPlot(test, group.by = c("orig.ident"), 
             shuffle = T, 
             raster = F, 
             label = F, 
             combine = T, 
             pt.size = 0.15,
             reduction = "ReUMAP_50")+ 
  theme_scp()+
  labs(x = "UMAP_1", y = "UMAP_2")+
  #NoLegend()+
  ggtitle("Subject")+
  scale_color_manual(values = as.vector(pal_subject_pal))+
  theme(axis.text = element_blank(), legend.direction = "vertical",   # Remove axis text
        axis.ticks = element_blank(),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 25, face = "bold"),
        legend.key.size = unit(2, "cm"),
        legend.position = "bottom")+
  NoLegend()+
  guides(color = guide_legend(override.aes = list(size = 16)))  

pdf(file.path(folder, "PPFE_IPF_EPI_subject_UMAP_RPCA1.pdf"), height = 14, width = 14)
p
dev.off()
```

## UMAP - Disease
```{r}
p <- DimPlot(test, group.by = c("disease.ident"),
          shuffle = T, raster = F, label = F, combine = T, pt.size = 0.15,
          reduction = "ReUMAP_50", 
        cols =dis_pal)+
  theme_scp()+ggtitle("Disease")+#NoLegend()+
  labs(x = "UMAP_1", y = "UMAP_2")+
  theme(axis.text = element_blank(), 
        legend.direction = "horizontal",# Remove axis text
        axis.ticks = element_blank(),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 25, face = "bold"),
        #legend.key.size = unit(6, "cm"),
        legend.position = "bottom")+
  guides(color = guide_legend(override.aes = list(size = 16)))  


pdf(file.path(folder, "PPFE_IPF_Lineage_Disease_UMAP_RPCA1_legend.pdf"), height = 14, width = 14)
p
dev.off()
```




```{r, fig.width=28, fig.height=14}
png(filename = file.path(folder, "Figure6_Disease_Cohort_UMAP.png"), height = 7, width = 14, units = "in", res = 1200)
#pdf(file = "PPFE_Mesenchyme_subject_Disease_UMAP.pdf", width = 14, height = 7)
p
dev.off()

pdf(file.path(folder, "Figure6_Disease_Cohort_UMAP.pdf"), height = 7, width = 14)
#pdf(file = "PPFE_Mesenchyme_subject_Disease_UMAP.pdf", width = 14, height = 7)
p
dev.off()
p
```

#==================
# Statistics

## Wilcox Rank Sum Test & Tukey-Test posthoc
```{r}
library(rstatix)
library(writexl)
summary_df <- ggData %>%
  group_by(Celltype_ID, Disease) %>%
  summarise(
    mean_value = mean(value, na.rm = TRUE),
    sd_value = sd(value, na.rm = TRUE)
  ) %>%
  ungroup()
write_xlsx(summary_df, file.path(folder, "PPFE_IPF_all_by_lineage_freq.xlsx"))

wilcox_results <- ggData %>%
  group_by(Celltype_ID) %>%
  wilcox_test(value ~ Disease) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj") %>%
  ungroup()

wilcox_results
write_xlsx(wilcox_results, file.path(folder, "PPFE_IPF_by_lineage_freq_wilcox.xlsx"))

final_df <- summary_df %>%
  left_join(wilcox_results, by = "Celltype_ID")

library(writexl)
write_xlsx(final_df, file.path(folder, "PPFE_IPF_allmerged_by_lineaeg_freq_Wilcox_rank_sum_test_merged.xlsx"))
```

```{r}
library(dplyr)
library(openxlsx)

# Initialize a workbook
wb <- createWorkbook()

# Germany_PPFE
median_frequencies_Germany_PPFE <- ggData %>%
  filter(Disease == "Germany_PPFE") %>%
  group_by(Celltype_ID) %>%
  summarize(
    median_freq = mean(value),
    sd_freq = sd(value)
  ) %>%
  arrange(-median_freq)

addWorksheet(wb, "Germany_PPFE")
writeData(wb, "Germany_PPFE", median_frequencies_Germany_PPFE)

# France_PPFE
median_frequencies_France_PPFE <- ggData %>%
  filter(Disease == "France_PPFE") %>%
  group_by(Celltype_ID) %>%
  summarize(
    median_freq = mean(value),
    sd_freq = sd(value)
  ) %>%
  arrange(-median_freq)

addWorksheet(wb, "France_PPFE")
writeData(wb, "France_PPFE", median_frequencies_France_PPFE)

# Germany_CTRL
median_frequencies_Germany_CTRL <- ggData %>%
  filter(Disease == "Germany_CTRL") %>%
  group_by(Celltype_ID) %>%
  summarize(
    median_freq = mean(value),
    sd_freq = sd(value)
  ) %>%
  arrange(-median_freq)

addWorksheet(wb, "Germany_CTRL")
writeData(wb, "Germany_CTRL", median_frequencies_Germany_CTRL)

# Belgian_IPF
median_frequencies_Dutch_IPF <- ggData %>%
  filter(Disease == "Belgian_IPF") %>%
  group_by(Celltype_ID) %>%
  summarize(
    median_freq = mean(value),
    sd_freq = sd(value)
  ) %>%
  arrange(-median_freq)

addWorksheet(wb, "Belgian_IPF")
writeData(wb, "Belgian_IPF", median_frequencies_Dutch_IPF)

# Save the workbook
saveWorkbook(wb, "PPFE_IPF_Celltype_Frequencies_ALL_b1b2_RPCA.xlsx", overwrite = TRUE)
```

#======================
# Build Heatmap

### Marker Gene Selection
```{r}
Lineage_genes <- c( "FXYD3", "EPCAM", "ELF3",#"EPCAM", "SLC34A2", "KRT7", "NKX2-1",
           "COL1A2", "DCN", "MFAP4", #"COL1A1", "COL3A1", "ELN", "COL6A2",
           "CLDN5", "CDH5", "CLEC14A", # "CDH5",  "PECAM1", "VWF", "BMPR2",
           "FCER1G", "C1orf162", "CLEC7A", #"CD68", "LYZ", "TLR2", "CSF1R",
           "CORO1A", "ISG20",  "PTPRC" #"CD2", "IL7R", "IL2RG", "CCL5"
           )
```

### Re-Order the Celltypes
```{r}
test$Celltype_ID_final_heat %>% unique()

test$Celltype_ID_final_heat <- factor(test$Celltype_ID_final, levels = c(
    "AEC2", "AEC1", "Club", "AEC_intermediate", "Basal", "Secretory", 
    "Aberrant_Basaloid", "Goblet", "Ciliated", "PNEC", "Mesothelium","Schwann_Neu",
   "Pericytes", "Alveolar_Fb", "Adventitial_Fb", "Subpleural_Fb", 
    "Adventitial_like_Fb", "Smooth_Muscle", "CTHRC1_MyoFb", "Airway_Fb",
   "EC_pulmonary_venous", "EC_arterial", "EC_general_cap", "EC_systemic_venous", 
    "EC_aerocyte", "EC_lymphatic",
   "Fibrotic_Macrophage", "Monocyte_derived_Macrophage", "Alveolar_Macrophage", 
    "Monocyte_nonClassical", "Monocyte_Classical", "DC2", "DC1", "Mast", 
    "IFNresp_Monocyte_derived_Macrophage", "pDC", "Langerhans-Cells", 
    "Macrophage_Cycling", "Neutro",
   "CD4+_TC_naive", "CD4+_TC_effector_memory", "CD4+_TC_TH1", "Treg", 
    "NK_Cells_CD16+", "CD8+_TC_resident_memory", "B_Cells_memory", 
    "CD8+_TC_effector_memory", "ILC3", "gamma_delta_TC", "CD4+_TC_TH17", 
    "CD4+_TC_Cytotoxic", "B_Cells_naive", "Plasma_Cells", "NK_Cells_CD16-")
)
```

### Start Heatmap-Script
```{r}
### with multithreading, first make all possible combinations
FullDataset_Object <-  test  

celltypes_to_plot <- levels(as.factor(test$Celltype_ID_final_heat))

FullDataset_Object$cellBarcode <- colnames(FullDataset_Object)

cellTypes <- levels(as.factor(FullDataset_Object$Celltype_ID_final_heat))

cellTypes <- cellTypes[cellTypes %in% celltypes_to_plot]

Lineages <- levels(test$Lineage)

meta.data.sub <- FullDataset_Object@meta.data[,c("Cohort",
                                                 "disease.ident", # cohort.ident condition
                                                 "Celltype_ID_final_heat", # cell.type.ident
                                                 "orig.ident_final", # subject.ident
                                                 "cellBarcode")]



get.CT.DS.subj.vector <- function(cellTypes){
  
  tmp.meta.data <- meta.data.sub %>% dplyr::filter(Celltype_ID_final_heat == cellTypes)
  
  cohorts <- unique(tmp.meta.data$Cohort)
  
  disease <- unique(tmp.meta.data$disease.ident)
  
  subjects <- unique(tmp.meta.data$orig.ident_final)
  
  tmp.CT.DS.subj.vector <- vector()
  
  for(j in 1:length(cohorts)){
    
    for(k in 1:length(disease)){
      
      for(l in 1:length(subjects)){
      
      temp.cells <- tmp.meta.data %>% dplyr::filter(Cohort==cohorts[j] & disease.ident==disease[k] & orig.ident_final==subjects[l]) %>% pull(cellBarcode)
      
      if ( length(temp.cells) >15 ) { # >=
        
        tmp.CT.DS.subj.vector <- c(tmp.CT.DS.subj.vector, paste(cellTypes, cohorts[j], disease[k], subjects[l], sep="__"))
        
        }
      }
      
    }

  }
  
  cat("Completed for ", cellTypes, ".\n", sep="")
  
  return(tmp.CT.DS.subj.vector)
  
}

celltype_cohort_subject.list <- parallel::mclapply(cellTypes, get.CT.DS.subj.vector, mc.cores=1)

celltype_cohort_subject <- unlist(celltype_cohort_subject.list)
celltype_cohort_subject %>% tail(300)
```
```{r}
library(Matrix)

DefaultAssay(FullDataset_Object) <- "RNA"

get.SubjectcohortCellTypeAvg <- function(celltype_cohort_subject){
  
  temp.cell.type <- strsplit(as.character(celltype_cohort_subject),"__")[[1]][1]
  
  temp.cohort <- strsplit(as.character(celltype_cohort_subject),"__")[[1]][2]
  
  temp.disease <- strsplit(as.character(celltype_cohort_subject),"__")[[1]][3]
  
  temp.subject <- strsplit(as.character(celltype_cohort_subject),"__")[[1]][4]
  
  temp.meta.data <- FullDataset_Object@meta.data[,c("Cohort", "disease.ident", "Celltype_ID_final_heat", "orig.ident_final", "cellBarcode")]
  
  temp.cells <- temp.meta.data %>% dplyr::filter(Celltype_ID_final_heat==temp.cell.type & Cohort==temp.cohort & disease.ident==temp.disease &
                                            
                                            orig.ident_final==temp.subject) %>% pull(cellBarcode)
  
  if (length(temp.cells) > 15) {
    
    tmp.df <- as.data.frame(rowMeans(GetAssayData(FullDataset_Object)[,temp.cells]))       
    
  } else {
    
    tmp.df <- as.data.frame(GetAssayData(FullDataset_Object)[,temp.cells])
    
    cat("Subject",temp.subject,"only has 1",temp.cell.type,"cell, using singlet for",temp.cohort,"representation...\n",sep=" ")
    
  }
  
  colnames(tmp.df) <- paste(celltype_cohort_subject)
  
  return(tmp.df)
  
}



collapsed.mtx.list <- parallel::mclapply(celltype_cohort_subject, get.SubjectcohortCellTypeAvg, mc.cores=1)

collapsed.SubjectcohortCellTypeAvg.mtx <- Matrix(as.matrix(do.call(cbind, collapsed.mtx.list)), sparse = TRUE)

dim(collapsed.SubjectcohortCellTypeAvg.mtx) # Average for each subject for each gene 

```
```{r}
heatmap_metadata <- as.data.frame(cbind(colnames(collapsed.SubjectcohortCellTypeAvg.mtx),
                                        
                                        sapply(strsplit(as.character(colnames(collapsed.SubjectcohortCellTypeAvg.mtx)),"__"), `[`, 1),
                                        
                                        sapply(strsplit(as.character(colnames(collapsed.SubjectcohortCellTypeAvg.mtx)),"__"), `[`, 2),
                                        
                                        sapply(strsplit(as.character(colnames(collapsed.SubjectcohortCellTypeAvg.mtx)),"__"), `[`, 3),
                                        
                                        sapply(strsplit(as.character(colnames(collapsed.SubjectcohortCellTypeAvg.mtx)),"__"), `[`, 4)
                                        )
                                  )


colnames(heatmap_metadata) <- c("cell.ident","cell.type.ident", "cohort.ident", "disease.ident", "subject.ident")


heatmap_metadata
```
```{r}
heatmap_metadata$cell.type.ident <- factor(heatmap_metadata$cell.type.ident, levels=celltypes_to_plot)

heatmap_metadata$cohort.ident <- factor(heatmap_metadata$cohort.ident, levels=c("Germany", "France","Belgium"))

heatmap_metadata$disease.ident <- factor(heatmap_metadata$disease.ident, levels=c( "CTRL", "PPFE", "IPF"))

heatmap_metadata
```
```{r}
cell_order <- heatmap_metadata %>% arrange(cell.type.ident, cohort.ident, disease.ident, subject.ident ) %>% pull(cell.ident)

cohort_order <- heatmap_metadata %>% arrange(cell.type.ident, cohort.ident,  disease.ident, subject.ident ) %>% pull(cohort.ident)

disease_order <- heatmap_metadata %>% arrange(cell.type.ident, cohort.ident, disease.ident, subject.ident ) %>% pull(disease.ident)

celltype_order <- heatmap_metadata %>% arrange(cell.type.ident, cohort.ident,  disease.ident, subject.ident ) %>% pull(cell.type.ident)

subject_order <- heatmap_metadata %>% arrange(cell.type.ident, cohort.ident,  disease.ident, subject.ident ) %>% pull(subject.ident )

celltype_order %>% levels()
```

```{r}
heatmap_df1 <-  as.matrix(collapsed.SubjectcohortCellTypeAvg.mtx[Lineage_genes,cell_order])
heatmap_df <- heatmap_df1
```

### Normalize by Cohort: Hannover vs. Leuven
```{r}
# Pull PPFE Cells
ppfe_cells <- heatmap_metadata %>% 
    dplyr::filter(cohort.ident %in% c("Germany", "France")) %>% 
    pull(cell.ident)

# Filter for IPF Cells
ipf_cells <- heatmap_metadata %>% dplyr::filter(cohort.ident == "Belgium") %>% pull(cell.ident)

# Merge Sub Dataframes
heatmap_df_ppfe <- heatmap_df[, ppfe_cells]
heatmap_df_ipf <- heatmap_df[, ipf_cells]


heatmap_metadata %>% dplyr::filter(cohort.ident == "Belgium")

# Assign Unity Normalization function
myUnityNormalize <- function(x){(x-min(x))/(max(x)-min(x))}

# Normalize each subset
heatmap_df_ppfe_normalized <- t(apply(heatmap_df_ppfe, MARGIN = 1, FUN = myUnityNormalize))
heatmap_df_ipf_normalized <- t(apply(heatmap_df_ipf, MARGIN = 1, FUN = myUnityNormalize))


# Combine both normalized datasets
heatmap_df_normalized <- cbind(heatmap_df_ppfe_normalized, heatmap_df_ipf_normalized)

# Get old order back
heatmap_df_normalized <- heatmap_df_normalized[,cell_order]  
```


### Create Ordering for column split
```{r, fig.width=20, fig.height=9.5}
Epithelial = c(
    "AEC2", "AEC1", "Club", "AEC_intermediate", "Basal", "Secretory", 
    "Aberrant_Basaloid", "Goblet", "Ciliated", "PNEC", "Mesothelium","Schwann_Neu"
  )
Mesenchymal = c(
    "Pericytes", "Alveolar_Fb", "Adventitial_Fb", "Subpleural_Fb", 
    "Adventitial_like_Fb", "Smooth_Muscle", "CTHRC1_MyoFb", "Airway_Fb"
  )
Endothelial = c(
    "EC_pulmonary_venous", "EC_arterial", "EC_general_cap", "EC_systemic_venous", 
    "EC_aerocyte", "EC_lymphatic"
  )
Myeloids = c(
    "Fibrotic_Macrophage", "Monocyte_derived_Macrophage", "Alveolar_Macrophage", 
    "Monocyte_nonClassical", "Monocyte_Classical", "DC2", "DC1", "Mast", 
    "IFNresp_Monocyte_derived_Macrophage", "pDC", "Langerhans-Cells", 
    "Macrophage_Cycling", "Neutro"
  )
Lymphocytes = c(
    "CD4+_TC_naive", "CD4+_TC_effector_memory", "CD4+_TC_TH1", "Treg", 
    "NK_Cells_CD16+", "CD8+_TC_resident_memory", "B_Cells_memory", 
    "CD8+_TC_effector_memory", "ILC3", "gamma_delta_TC", "CD4+_TC_TH17", 
    "CD4+_TC_Cytotoxic", "B_Cells_naive", "Plasma_Cells", "NK_Cells_CD16-"
  )

split_order <- celltype_order 

# Function to rename based on occurrence in groups
rename_celltypes <- function(celltypes, Epithelial, Mesenchymal, Endothelial, Myeloids, Lymphocytes) {
  sapply(celltypes, function(ct) {
    if (ct %in% Epithelial) {
      return("Epithelium")
    } else if (ct %in% Mesenchymal) {
      return("Mesenchyme")
    } else if (ct %in% Endothelial) {
      return("Endothelium")
    } else if (ct %in% Myeloids) {
      return("Myeloids")
    } else if (ct %in% Lymphocytes) {
      return("Lymphocytes")
    } else {
      return("Unknown")
    }
  })
}

# Rename cell types
renamed_celltypes <- rename_celltypes(celltype_order, Epithelial, Mesenchymal, Endothelial, Myeloids, Lymphocytes)

# View renamed cell types
renamed_celltypes %>% unique()

renamed_celltypes <- factor(renamed_celltypes, levels = c("Epithelium", "Mesenchyme", "Endothelium", "Myeloids", "Lymphocytes"))

# Check final ordering of factor levels
renamed_celltypes %>% levels()
```


```{r}
library(ComplexHeatmap)
#dis_pal <- list("royalblue","red3", "darkgreen")
#cohort_pal <- list("darkgrey", "lightgrey")
cohort_pal_heat <- palette_scp(x = test$Cohort, palcolor = c("grey88", "grey58", "grey18"))
cohort_colors <- palette_scp(x = cohort_order, palcolor = cohort_pal_heat, n = 3, matched = F)
disease_colors <- palette_scp(x = disease_order, palcolor = pal_dis_pal, n = 3, matched = F) 
celltype_colors <- palette_scp(x = celltype_order, palcolor = pal_cluster_scp_pal, matched = F)
subject_colors <- SCP::palette_scp(x = subject_order,  palcolor = pal_subject_pal, n = 111, matched = F)
lineage_colors <- SCP::palette_scp(x = renamed_celltypes, palcolor = lin_pal, n = 5, matched = F)


# Plot Annotation without with legend
heatmap_cohort_annotation <- HeatmapAnnotation(lineage = renamed_celltypes, cell_type=celltype_order, cohort=cohort_order, disease = disease_order, subject=subject_order, col = list(cohort=cohort_colors, disease = disease_colors, lineage = lineage_colors, cell_type=celltype_colors, subject=subject_colors), show_legend = c("subject" = FALSE))

# Plot Annotation without any legend
heatmap_cohort_annotation <- HeatmapAnnotation(lineage = renamed_celltypes, 
                                               cell_type=celltype_order,
                                               cohort=cohort_order, 
                                               disease = disease_order, 
                                               subject=subject_order, 
                                               col = list(cohort=cohort_colors, 
                                                          disease = disease_colors, 
                                                          lineage = lineage_colors, 
                                                          cell_type=celltype_colors, 
                                                          subject=subject_colors),
                                               show_legend = F)#c("subject" = FALSE))

```

```{r, fig.width=20, fig.height=9.5}
# Plot the final Heatmap
library(viridis)
p <- Heatmap(heatmap_df_normalized, name = "Scaled Expression",
        col = viridis(256), 
        cluster_rows = F, 
        show_heatmap_legend = F,
        cluster_columns = F, 
        show_column_names = FALSE, 
        top_annotation=heatmap_cohort_annotation, 
        column_split=renamed_celltypes, 
        #row_split = c(rep("1. Celltype Marker", nrow(heatmap_df1)), rep("2. Aberrant Basaloid Marker", nrow(heatmap_df2))),  # split to create visual separation
       # row_gap = unit(c(2, 2), "mm"),  # larger gap around the separation line
        column_title = NULL, 
        use_raster=FALSE,
        column_gap = unit(0.5, "mm"))
p
```



```{r, fig.width=20, fig.height=9.5}
pdf(file.path(folder, "PPFE_IPF_ALL_LINEAGES_HEATMAP_RPCA_Aberrant_Features_bycohort_recol.pdf"),
    height = 7, width = 3.0)
p
dev.off()


pdf(file = file.path(folder, "PPFE_IPF_ALL_LINEAGES_HEATMAP_RPCA_Aberrant_Features_bycohort_legend.pdf"), 
    height = 9.5, width = 20)
p
dev.off()

```


#==========================

## Figure S7 - Correlation Plot
```{r}
library(ggplot2)
library(dplyr)
library(ggpubr)
lineages <- merged_df$Celltype_ID %>% levels()

# Define custom colors and shapes
disease_colors <- c("Germany_PPFE" = "darkred", "France_PPFE" = "pink")
disease_shapes <- c("Germany_PPFE" = 16, "France_PPFE" = 17)  # e.g., filled circle and triangle

# Generate the plots
plot_list <- lapply(lineages, function(x){
  cor_data <- merged_df %>% 
    filter(Celltype_ID == x) %>%
    filter(Disease %in% c("Germany_PPFE", "France_PPFE")) %>%
    mutate(Frequency = value * 100,
           Remodeled_Parenchyma = Remodeled_Alv_Ratio * 100)

  ggplot(cor_data, aes(x = Frequency, y = Remodeled_Parenchyma)) +
    geom_point(aes(color = Disease, shape = Disease), size = 2) +  # color & shape for points only
    geom_smooth(method = "lm", se = TRUE, color = "black") +       # single correlation line
    stat_cor(method = "pearson") +
    scale_color_manual(values = disease_colors) +
    scale_shape_manual(values = disease_shapes) +
    labs(
      x = paste0("Frequency of ", x, " (%)"),
      y = "Remodeled Parenchyma (%)",
      title = paste0("Correlation: ", x)
    ) +
    theme_scp()
})

# Arrange and save
combined_plot <- ggarrange(plotlist = plot_list, ncol = 2, nrow = ceiling(length(plot_list)/2), common.legend = T)
ggsave("Immune_Correlation_Multipanel.pdf", combined_plot, width = 6, height = 8)

```

# Save Final Dataset
```{r}
saveRDS(test, "PPFE_Germany_France_CTRL_IPF.int.11032025.rds")
```