---
title: "PPFE_IPF_Mes_Enrichment"
author: "Jannik_Ru"
date: "2025-01-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      fig.width = 14,
                      fig.height = 7,
                      #dpi = 1200,
                      dev = "png",
                      cache = F)
```

# Prequisite
```{r}
# Adventitial-like fibroblasts were later labeled as Elstofibrotic Fibroblasts in the Manuscript
```

# Load Packages
```{r}
library(SCP)
library(Seurat)
library(ggplot2)
library(ggplot2)
library(matrixStats)
library(dplyr)
library(tidyseurat)
library(readxl)
library(writexl)
library(SeuratWrappers)
```

```{r}
PPFE_mes <- readRDS("./PPFE_Mes_Lineage.rds") 
```

#=============

# Enrichment Analyse

## GOA
```{r, fig.width=14}
# Remove IPF and CTRL
PPFE_mes <- subset(PPFE_mes, disease.ident == "PPFE")

# Run DE Test based on MAST
PPFE_mes  <- RunDEtest(srt = PPFE_mes, 
                       group_by = "Celltype_ID_final_coarse", 
                       fc.threshold = 1.5, 
                       only.pos = FALSE, 
                       test.use = "MAST")
# Run GOA Analysis
PPFE_mes  <- RunEnrichment(
  srt = PPFE_mes , 
  group_by = "Celltype_ID_final_coarse", 
  db = c("KEGG", "WikiPathway", "Reactome","GO_BP", "MSigDB") , 
  species = "Homo_sapiens", 
  DE_threshold = "avg_log2FC > log2(1.5) & p_val_adj < 0.05",
  test.use = "MAST") # Get DE Genes from MAST DEG Slot in the tools subdir of S4 Object

# Assign Color Palette
SCP::palette_scp(palcolor = c("#e31a1c", "#fb9a99","#cab2d6","#6a3d9a"),
                 x =  list("Subpleural_Fb", "Adventitial_Fb","Adventitial_like_Fb","CTHRC1_MyoFb" , n = 4), matched = T) -> palcol
```

#=============

## Barplots 

### Subpleural Fb
```{r, fig.height=14, fig.width=10}
p <- EnrichmentPlot(
  srt = PPFE_mes , group_by = "Celltype_ID_final_coarse", group_use = c("Subpleural_Fb"), 
  test.use = "MAST", 
  color_by = "Groups", 
  palcolor = "#fb9a99", 
  plot_type = "bar", 
  topTerm = 15,  
  db = c("Reactome"), # Filter for Reactome
  combine = T, 
  ncol = 1, nrow = 4, lineheight = 0.9, character_width = 90, 
)
p

# Export Plot
png("PPFE_mes_EnrichBarPlot_map_GOA_Reactome_SubleuralFb.png", res = 1200, units = "in", width = 14, height = 6)
p
dev.off()
```

### Adventitial_Fb
```{r}
# Enrichment Plot
p <- EnrichmentPlot(
  srt = PPFE_mes , group_by = "Celltype_ID_final_coarse", 
  group_use = c("Adventitial_Fb"), 
  test.use = "MAST", 
  color_by = "Groups", 
  palcolor = "#6a3d9a", 
  plot_type = "bar", 
  topTerm = 15,  
  db = c("Reactome"), # Filter for Reactome
  combine = T, 
  ncol = 1, nrow = 4, lineheight = 0.9, character_width = 90, 
)
p

# Export as PNG 
png("PPFE_mes_EnrichBarPlot_map_GOA_Reactome_AdvntlFb.png", 
    res = 1200, units = "in", 
    width = 14, height = 6)
dev.off()
```

### Adventitial-like Fb | Elastofibroblast Fb
```{r}
p <- EnrichmentPlot(
  srt = PPFE_mes , 
  group_by = "Celltype_ID_final_coarse", 
  group_use = c("Adventitial_like_Fb"), 
  test.use = "MAST", 
  color_by = "Groups", 
  palcolor = "#cab2d6", 
  plot_type = "bar", 
  topTerm = 15,  
  db = c("Reactome"), 
  combine = T, 
  ncol = 1, nrow = 4, lineheight = 0.9, character_width = 90, 
)
p

# Export as PNG 
png("PPFE_mes_EnrichBarPlot_map_GOA_Reactome_Advntl_likeFb.png", 
    res = 1200, units = "in", width = 14, height = 6)
p
dev.off()
```

### CTHRC1_fibrotic_Fb
```{r}
p <- EnrichmentPlot(
  srt = PPFE_mes , group_by = "Celltype_ID_final_coarse", group_use = c("CTHRC1_MyoFb"), 
  test.use = "MAST", 
  color_by = "Groups", 
  palcolor = "#e31a1c", 
  plot_type = "bar", 
  topTerm = 15,  
  db = c("Reactome"), 
  combine = T, 
  ncol = 1, nrow = 4, lineheight = 0.9, character_width = 90, 
)
p

png("PPFE_mes_EnrichBarPlot_map_GOA_Reactome_MyoFb.png", res = 1200, units = "in", width = 14, height = 6)
p
dev.off()
```

#==============
## Enrichmap-Plots (Figure S4)

### GOA Enrichmap Plot - Adventitial-like Fb | Elastofibroblast Fb
```{r, fig.width=14, fig.height=14}
# Calculate Enrichmap
p <- EnrichmentPlot(
  srt = PPFE_mes, group_by = "Celltype_ID_final_coarse", group_use = "Adventitial_like_Fb", db = "Reactome",
  enrichmap_labelsize = 10,
  plot_type = "enrichmap", enrlichmap_nlabel = 10, topTerm = 30)+
  Seurat::DarkTheme()+ theme( panel.background = element_rect(fill = "white"),
                      axis.ticks = element_blank(),
                      axis.text = element_blank())

# Export greedy clusters from plot 
p[["layers"]][[1]][["data"]] -> res_enrich_GOA_clusetred
write_xlsx(res_enrich_GOA_clusetred, "res_enrich_GOA_clustered_Adventitial_like_Fb.xlsx")

# Export Plot
png("PPFE_mes_Enrichmap_map_GOA_Reactome_Adv_like.png", res = 300, units = "in", width = 18, height = 10)
p
dev.off()

pdf(file = "PPFE_mes_Enrichmap_map_GOA_Reactome_Adv_like.pdf", width = 26, height = 18)
p
dev.off()
```

#### Enrichmap - Cluster C4 as Network
```{r}
# Filter the GOA Dataframe according to Cluster of interst 
res_enrich_GOA_clusetred %>% filter(clusters == "C3") -> filtered_df

# Export original Enrichment Results from Seurat Object ! 
PPFE_mes@tools$Enrichment_Celltype_ID_final_coarse_wilcox -> enrich
enrich_enrichment <- enrich$enrichment

# Filter orinigal enrichment dataframe for Genset IDs occuring in the Cluster of interst from the enrichmap 
enrich$enrichment <- enrich_enrichment[enrich_enrichment$ID %in% filtered_df$name,]

# Plot the Network Plot per Cluster 
pc4 <- EnrichmentPlot(
  res = enrich,  
  group_use = "Adventitial_like_Fb", 
  db = "Reactome",
  plot_type = "network", 
  enrlichmap_nlabel = 8, 
  topTerm = 300)+
  DarkTheme()+ 
  theme(axis.line = element_line(colour = "black"),
                      plot.title = ggplot2::element_text(size = 20, face = "bold"),
                      panel.grid.major = element_blank(),
                      panel.grid.minor = element_blank(),
                      panel.border = element_blank(),
                      panel.background = element_blank(), 
                      axis.text = element_blank(),   # Remove axis text
                      axis.ticks = element_blank(),
                      strip.text = element_text(size = 16, face = "bold", colour = "white"),
                      strip.background = element_blank(),  # Remove facet background
                      panel.spacing = unit(1, "lines"),  # Remove space between facets
                      plot.margin = margin(0, 0, 0, 0) 
                      )
png("PPFE_mes_Enrichmap_GOA_Reactome_Adv_like_fb_C4.png", res = 600, units = "in", width = 18, height = 10)
pc4
dev.off()

pdf(file = "PPFE_mes_Enrichmap_GOA_Reactome_Adv_like_fb_C4.pdf",width = 10, height = 6, bg = "black")
pc4
dev.off()
```


### GOA Enrichmap Plot - Adventitial_Fb
```{r, fig.width=14}
p <- EnrichmentPlot(
  srt = PPFE_mes, group_by = "Celltype_ID_final_coarse", group_use = "Adventitial_Fb", db = "Reactome", enrichmap_labelsize = 9,
  plot_type = "enrichmap", enrlichmap_nlabel = 50, topTerm = 50#
)+DarkTheme()+ theme( panel.background = element_rect(fill = "white"),
                      axis.ticks = element_blank(),
                      axis.text = element_blank())

# Export greedy clusters from plot 
p[["layers"]][[1]][["data"]] -> res_enrich_GOA_clusetred
write_xlsx(res_enrich_GOA_clusetred, "res_enrich_GOA_clustered_Adventitial_Fb.xlsx")

# Plot
png("PPFE_mes_Enrichmap_map_GOA_Reactome_AdvFb.png", res = 300, units = "in", width = 18, height = 10)
p
dev.off()

pdf(file = "PPFE_mes_Enrichmap_map_GOA_Reactome_AdvFb.pdf", width = 22, height = 14)
p
dev.off()
```

#### Enrichmap Cluster C1 as network plot (Figure 2K)
```{r, fig.width=10, fig.height=10}
# Filter the GOA Dataframe according to Cluster of interst 
res_enrich_GOA_clusetred %>% filter(clusters == "C1") -> filtered_df

# Export original Enrichment Results from Seurat Object ! CAVE Don't use simple enrichment dataframe ! 
PPFE_mes@tools$Enrichment_Celltype_ID_final_coarse_wilcox -> enrich
enrich_enrichment <- enrich$enrichment

# Filter orinigal enrichment dataframe for Genset IDs occuring in the Cluster of interst from the enrichmap 
enrich$enrichment <- enrich_enrichment[enrich_enrichment$ID %in% filtered_df$name,]

# Plot the Network Plot per Cluster 
pC1 <- EnrichmentPlot(
  res = enrich,  group_use = "Adventitial_Fb", db = "Reactome",#group_by = "Celltype_ID_final_coarse",
  plot_type = "network", 
  enrlichmap_nlabel = 9, 
  enrichmap_labelsize = 10,
  #legend.position = "bottom",
  topTerm = 300#, enrichmap_label = "feature"
  )+DarkTheme()+ theme(axis.line = element_line(colour = "black"),
                      plot.title = ggplot2::element_text(size = 20, face = "bold"),
                      panel.grid.major = element_blank(),
                      panel.grid.minor = element_blank(),
                      panel.border = element_blank(),
                      panel.background = element_blank(), 
                      axis.text = element_blank(),   # Remove axis text
                      axis.ticks = element_blank(),
                      strip.text = element_text(size = 16, face = "bold", colour = "white"),
                      strip.background = element_blank(),  # Remove facet background
                      panel.spacing = unit(1, "lines"),  # Remove space between facets
                      plot.margin = margin(0, 0, 0, 0) 
                      )
pdf(file = "PPFE_mes_Enrichmap_GOA_Reactome_Adv_fb_C1.pdf", width = 11, height = 7)
pC1
dev.off()

png("PPFE_mes_Enrichmap_GOA_Reactome_Adv_fb_C1.png", res = 600, units = "in", width = 10, height = 10)
pC1
dev.off()
```

#=============
# save Enrichment data
```{r}
library(writexl)
write_xlsx(PPFE_mes@tools$Enrichment_Celltype_ID_final_coarse_MAST$enrichment, "PPFE_Mes_MAST_Enrichment.xlsx")
```